### 2022-11-27 methods-and-results.Rmd
:sparkles: descr tbl

---
output: latex_fragment
editor_options:
  chunk_output_type: console
bibliography: references.bib
csl: apa.csl
---

```{r setup, include=FALSE}
# R Studio Clean-Up
cat("\014") # clear console
# rm(list=ls()) # clear workspace - use restart R instead [cmd/alt + shift + F10]
gc() # garbage collector

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  #"tidyverse",
  "MASS",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  "haven",
  "RColorBrewer",
  "plotly",
  "grid",
  "gridExtra",
  "ggpattern",
  "lme4",
  "nlme",
  "jtools",
  "gtsummary",
  "sessioninfo",
  "tibble",
  "pander",
  "devtools",
  "mada",
  "data.table",
  "plyr",
  "dplyr",
  "tidyr",
  "Hmisc",
  "kableExtra",
  "papaja",
  "stringr",
  "stringi",
  "reshape2",
  "anytime",
  "lubridate",
  "purrr",
  "metafor",
  "dygraphs",
  "readxl", 
  "reshape",
  "factoextra",
  "Amelia", 
  "ThreeWay",
  "DescTools",
  # GAM
  "gratia",
  #"MASS",
  "mgcv",
  # For moving window models
  "Kendall",
  # For the Hamed & Rao correction of tau
  "modifiedmk", 
  "mgcViz",
  "shiny"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
anytime::addFormats("%Y-%m-%d %H:%M:%S")
anytime::addFormats("%d-%m-%Y")  
#options(rgl.useNULL = TRUE)

# Load Custom Packages
source("./scripts/functions/fun.panel.R")
source("./scripts/functions/themes.R")
source("./scripts/functions/binaryCor.R")
source("./scripts/functions/MlCorMat.R")
source("./scripts/functions/MlTbl.R")
source("./scripts/functions/metaLmer.R")
source("./scripts/functions/meanViz.R")
source("./scripts/functions/gam.R")
source("./scripts/functions/tsFeatureExtractor.R")

# Markdown Options
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # set working directory
knitr::opts_knit$get("root.dir") # check working directory
options(
  scipen = 999,
  digits = 4,
  width = 400
) # removes scientific quotation
# knitr::opts_chunk$set(echo = TRUE, cache = F, cache.path = rprojroot::find_rstudio_root_file('cache/')) # cache settings

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "Figures/",
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r import data, include=FALSE}
# Load variable name lookup table
varNames <- readxl::read_excel("preregistration/varNames.xlsx")

# import full datasets
load("data/s123Full.RData")

# import reduced datasets
load("data/s123Red.RData")

# import pre-calculated features for faster renders
load("data/features/featOut.RData")
```

```{r Preparation of Variable Names, include=FALSE}
varNamIndicesS1 <- c(
  "autonomy",
  "competence",
  "relatedness",
  "education_level.pre",
  "associationMerged.pre",
  "assimilation.pre",
  "separation.pre",
  "integration.pre",
  "marginalization.pre",
  "VIA_heritage.pre",
  "VIA_Dutch.pre",
  "SSAS_surrounding.pre",
  "SSAS_privat.pre",
  "SSAS_public.pre",
  "assimilation.post",
  "separation.post",
  "integration.post",
  "marginalization.post",
  "VIA_heritage.post",
  "VIA_Dutch.post",
  "rosenberg.post",
  "social_support.post",
  "stress.post",
  "discrimination.post",
  "discrimination_month.post",
  "NLE_1month.post",
  "NLE_6month.post",
  "NLE_12month.post"
  )

varNamS123Aux <- varNames %>%
  filter(
    aux != 0,
    studyS1 == "S1",
    studyS2 == "S2",
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull
varNamS1Aux <- varNames %>%
  filter(
    aux != 0,
    studyS1 == "S1"
  ) %>%
  select(varNam) %>%
  pull
varNamS2Aux <- varNames %>%
  filter(
    aux != 0,
    studyS2 == "S2"
  ) %>%
  select(varNam) %>%
  pull
varNamS3Aux <- varNames %>%
  filter(
    aux != 0,
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull

varNamS123PCA <- varNames %>%
  filter(
    pca != 0,
    studyS1 == "S1",
    studyS2 == "S2",
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull
varNamS1PCA <- varNames %>%
  filter(
    pca != 0,
    studyS1 == "S1"
  ) %>%
  select(varNam) %>%
  pull
varNamS2PCA <- varNames %>%
  filter(
    pca != 0,
    studyS2 == "S2"
  ) %>%
  select(varNam) %>% 
  pull
varNamS3PCA <- varNames %>%
  filter(
    pca != 0,
    studyS3 == "S3"
  ) %>%
  select(varNam) %>%
  pull

varNamOut <- c(
  "ResponseId",
  "relatednessNoInteraction",
  "relatednessSelf",
  "relatednessOther",
  "autonomy_Int",
  "autonomy_NoInt",
  "competence_Int",
  "competence_NoInt",
  varNamS1PCA[grepl('^MDMQ', varNamS1PCA)],
  varNamS2PCA[grepl('^ProSo|^AntiSo', varNamS2PCA)],
  varNamS3PCA[grepl(
    '^ProSo|^AntiSo|^agency|^autoFrust|^autoSat|^relatFrust|^relatSat|^compFrust|^compSat|^lonely[0-9]|^emotRegPos|^emotRegNeg',
    varNamS3PCA
  )]
)
varNamCore <- c(
  "PID",
  "TID",
  "TIDnum"
)
varNamNewAll <- c(
  "relatedness",
  "autonomy",
  "competence"
)
varNamNewS1 <- c(
  "relatedness",
  "autonomy",
  "competence",
  "alertness", 
  "calmness", 
  "valence"
)
varNamNewS2 <- c(
  "relatedness",
  "autonomy",
  "competence",
  "ProSo", 
  "AntiSo"
)
varNamNewS3 <- c(
  "relatedness",
  "autonomy",
  "competence",
  "ProSo", 
  "AntiSo",
  "agency",
  "autoFrust",
  "autoSat",
  "relatFrust",
  "relatSat",
  "compFrust",
  "compSat",
  "lonely",
  "emotRegPos",
  "emotRegNeg"
)

varNamNewCat <- c(
  "closeness_Calc",
  "gender_Calc",
  "ethnicity_Calc",
  "relationship_Calc"
)

varNamIntDep <- c(
  varNamS123PCA[grepl('^InteractionContext|AttitudesPartner|KeyNeedDueToPartner|^quality', varNamS123PCA)]
)

varNamIndiceItemsS1 <- varNames %>%
  filter(
    aux == -1,
    studyS1 == "S1"
  ) %>%
  select(varNam) %>%
  pull %>%
  append(., gsub(".pre|.post|_calc", "", varNamIndicesS1)) %>%
  unique

varNamS123MI <- c(varNamCore, varNamNewAll, varNamS123Aux)
varNamS1MI <- c(varNamCore, varNamIndicesS1, varNamS1Aux[!varNamS1Aux %in% varNamIndiceItemsS1]) #
varNamS2MI <- c(varNamCore, varNamNewS2, varNamS2Aux)
varNamS3MI <- c(varNamCore, varNamNewS3, varNamS3Aux)

varNamS123PCA <- c(varNamCore, varNamS123PCA[!varNamS123PCA %in% varNamOut])
varNamS1PCA <- c(varNamCore, varNamNewS1, varNamS1PCA[!varNamS1PCA %in% varNamOut])
varNamS2PCA <- c(varNamCore, varNamNewS2, varNamS2PCA[!varNamS2PCA %in% varNamOut])
varNamS3PCA <- c(varNamCore, varNamNewS3, varNamS3PCA[!varNamS3PCA %in% varNamOut])

varNamS123MIPsbl <- c(
  paste(varNamS123MI, rep(".pre", length(varNamS123MI)), sep = ""),
  varNamS123MI,
  paste(varNamS123MI, rep(".post", length(varNamS123MI)), sep = "")
)

varNamS123MiRed <- Reduce(
  intersect,
  list(
    dtS1Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
    dtS2Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names,
    dtS3Red %>% select(PID, TID, TIDnum, any_of(varNamS123MIPsbl)) %>% names
  )
)

idVars <- c("ID", "PID", "TID", "TIDnum", "date", "week", "study")
```

```{r variable names for all analyses, include=FALSE}
# ID Variables (just to re-iterate)
idVars <- idVars

# Common variables across studies 
varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]

# Non-Interaction Specific Variables
varNamS123NoInt <- varNames %>%
  filter(
    varNam %in% varNamS123PCA,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# Interaction Specific Variables
varNamS123Int <- varNames %>%
  filter(
    varNam %in% varNamS123PCA,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S1 variables
varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]

# S1 Non-Interaction Specific Variables
varNamS1NoInt <- varNames %>%
  filter(
    varNam %in% varNamS1Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S1 Interaction Specific Variables
varNamS1Int <- varNames %>%
  filter(
    varNam %in% varNamS1Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S2 variables
varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]

# S2 Non-Interaction Specific Variables
varNamS2NoInt <- varNames %>%
  filter(
    varNam %in% varNamS2Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S2 Interaction Specific Variables
varNamS2Int <- varNames %>%
  filter(
    varNam %in% varNamS2Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S3 variables
varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]

# S3 Non-Interaction Specific Variables
varNamS3NoInt <- varNames %>%
  filter(
    varNam %in% varNamS3Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S3 Interaction Specific Variables
varNamS3Int <- varNames %>%
  filter(
    varNam %in% varNamS3Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull
```

```{r variable names by response type and study, include=FALSE}
varNames %>%
  filter(
    pca != 0,
    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
  ) %>%
  mutate(
    contactSpecific = recode(
      .$contactSpecific,
      `-1` = "No Interaction Only",
      `0` = "unspecific",
      `1` = "Interaction Only"
    ),
    study1 = ifelse(!is.na(studyS1), 1, 0),
    study2 = ifelse(!is.na(studyS2), 1, 0),
    study3 = ifelse(!is.na(studyS3), 1, 0)
  ) %>% 
  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
  ungroup %>%
  t %>%
  as.data.frame %>%
  janitor::row_to_names(1) %>%
  replace(is.na(.), 0) %>%
  tibble::rownames_to_column(., var = "variable") %>%
  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
  mutate(across(c(study1, study2, study3), as.numeric)) %>%
  arrange(concept, unspecific, `Interaction Only`) %>% 
  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
  relocate(starts_with("study"), .after = last_col()) %>%
  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "l", rep("c", ncol(.)-2)),
      digits=2,
      caption = "Variables by Interaction Types and Study Availability") %>%
  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

```{r variables included in main analysis by response type, include=FALSE}
varNames %>%
  filter(
    pca != 0,
    varNam %in% c(varNamS123PCA, varNamS1PCA, varNamS2PCA, varNamS3PCA)
  ) %>%
  mutate(
    contactSpecific = recode(
      .$contactSpecific,
      `-1` = "No Interaction Only",
      `0` = "unspecific",
      `1` = "Interaction Only"
    ),
    study1 = ifelse(!is.na(studyS1), 1, 0),
    study2 = ifelse(!is.na(studyS2), 1, 0),
    study3 = ifelse(!is.na(studyS3), 1, 0)
  ) %>% 
  select(varNam, varGroup, contactSpecific, study1, study2, study3) %>%
  filter(study1 == 1 & study2 == 1 & study3 == 1) %>%
  group_by(varGroup, varNam, contactSpecific, study1, study2, study3) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = c("varGroup", "varNam", "study1", "study2", "study3"), values_from = n, names_sep=" - ") %>%
  ungroup %>%
  t %>%
  as.data.frame %>%
  janitor::row_to_names(1) %>%
  replace(is.na(.), 0) %>%
  tibble::rownames_to_column(., var = "variable") %>%
  separate(variable, c("concept", "variable", "study1", "study2", "study3"), " - ") %>%
  mutate(across(c(study1, study2, study3), as.numeric)) %>%
  #select(-study1, -study2, -study3) %>%
  arrange(concept, unspecific, `Interaction Only`) %>% 
  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
  relocate(starts_with("study"), .after = last_col()) %>%
  rename_with(., ~ gsub("study", "Study ", .x, fixed = TRUE)) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "l", rep("c", ncol(.)-2)),
      digits=2,
      caption = "Variables by Interaction Types and Study Availability") %>%
  add_header_above(c(" " = 2, "Contact Type" = 2, "Study Availability" = 3)) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

```{r missingness filtering, include=FALSE}
missInfo <- function(full = NA, reduced = NA){
  missTab <- data.frame(
    nFull = nrow(full),
    nRed = nrow(reduced),
    pptFull = length(unique(full$PID)),
    pptRed = length(unique(reduced$PID)),
    timeFull = length(unique(full$TID)),
    timeRed = length(unique(reduced$TID))
  ) %>%
    mutate(
      nDif = nFull-nRed,
      nDifPerc = nDif/nFull*100,
      pptDif = pptFull-pptRed,
      pptDifPerc = pptDif/pptFull*100,
      timeDif = timeFull-timeRed,
      timeDifPerc = timeDif/timeFull*100
    ) %>%
    select(
      nFull, nRed, nDif, nDifPerc,
      pptFull, pptRed, pptDif, pptDifPerc,
      timeFull, timeRed, timeDif, timeDifPerc
    )
    missTab
}

missS1 <- missInfo(full = dtS1$full, reduced = dtS1Red)
missS2 <- missInfo(full = dtS2$full, reduced = dtS2Red)
missS3 <- missInfo(full = dtS3$full, reduced = dtS3Red)

missS123 <- rbind(
  missS1,
  missS2,
  missS3
) %>%
  mutate(study = c(1, 2, 3)) %>%
  select(study, everything())

missS123 %>%
  kbl(.,
      escape = FALSE,
      booktabs = TRUE,
      align = "c", #c("l", rep("c", ncol(.)-1)),
      col.names = c("Study", rep(c("Full","Reduced","$\\Delta$", "%"), 3)),
      digits=2,
      caption = "Missingness Info by Study") %>%
  add_header_above(c(" " = 1, "Measurements" = 4, "Participants" = 4, "Timepoints" = 4)) %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )
```

\subsection{Data}
We used a data set following migration experiences collected by \citet[][]{Kreienkamp2022b}. The data set consists of three studies that followed migrants who had recently arrived in the Netherlands in their daily interactions with the Dutch majority group members. After a general migration-focused pre-questionnaire, participants were invited twice per day to report on their (potential) interactions with majority group members for at least 30 days. The short ESM surveys were sent out at around lunch (12pm) and dinner time (7pm). After the 30 day study period participants filled in a post-questionnaire that mirrored the pre-questionnaire. Participants received either monetary compensation or partial course credits based on the number of surveys they completed. For our empirical example we focus on the variables that were collected during the ESM surveys and were available in all three studies. Full methodological details are available in the empirical article by \citet[][]{Kreienkamp2022b}.

\subsubsection{Sample}
The original studies included `r sum(missS123$pptFull)` participants ($N_{S1}=$ `r missS123$pptFull[missS123$study==1]`, $N_{S2}=$ `r missS123$pptFull[missS123$study==2]`, $N_{S3}=$ `r missS123$pptFull[missS123$study==3]`) with a total of `r format(sum(missS123$nFull), big.mark=",")` measurements. The studies differed substantially in the maximum length of participation ($\text{max}(t_{S1})=$ `r missS123$timeFull[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeFull[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeFull[missS123$study==3]`). This was likely due to the option to continue participation without compensation in the later two studies. To make the three studies comparable in participation and time frames, we iteratively removed all measurement occasions and participants that had more than 45\% missingness \citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}. This procedure let to a final sample of `r sum(missS123$pptRed)` participants, who jointly produced `r format(sum(missS123$nRed), big.mark=",")` measurements. Importantly, both the participant repsonse-patterns and the time frame were now a lot more comparable ($\text{max}(t_{S1})=$ `r missS123$timeRed[missS123$study==1]`, $\text{max}(t_{S2})=$ `r missS123$timeRed[missS123$study==2]`, $\text{max}(t_{S3})=$ `r missS123$timeRed[missS123$study==3]`). For a full overview of the data reduction procedure and study-specific exclusions see Online Supplemental Material A.

\subsubsection{Variables}
Included 12 variables that were available across all three studies and captured information about the participant's interactions, as well as cognitive-, emotional-, and motivational self in relationship with the majority group. We chose these two aspects in particular because (1) the interaction-specific information exemplified the structural missingness issue of modern ESM data and (2) the motivational, emotional, and cognitive experience offered a diverse conceptualization of migration experience (beyond behavioral measurements) that is becoming more common in the literature \citep[][]{Kreienkamp2022d}. Full methodological details are available in Online Supplemental Material A, but basic item information, descriptives, and correlations are available in \tblref{tab:descrWide}.

```{r descrWide, include=FALSE}
dtAll <- rbind(
  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
    mutate(across(!TID & !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(ID,
         PID,
         study,
         date,
         week,
         TIDnum,
         any_of(varNamS123PCA)) %>%
  arrange(ID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

allMlCor <-
  MlCorMat(
    data = dtAll,
    id = "ID",
    selection = c(
      "InteractionContextAccidental", 
      "InteractionContextvoluntary", 
      "InteractionContextCooperative",
      "InteractionContextRepresentativeNL",
      "qualityMeaning",
      "qualityOverall", 
      "KeyNeedFulfillment", 
      "KeyNeedDueToPartner", 
      "AttitudesPartner",
      "DaytimeNeedFulfillment",
      "AttitudesDutch",
      "exWB"
    ),
    labels = c(
      "Interaction Accidental", 
      "Interaction Voluntary", 
      "Interaction Cooperative",
      "Interaction Representative NL",
      "Quality Meaningful",
      "Quality Overall", 
      "Core Need", 
      "Core Need Due to Partner", 
      "Attitudes Partner",
      "Daytime Core Need",
      "Attitudes Dutch",
      "Well-being"
    )
  )

# Sample Descriptives Table --- wider version
allMlCor %>%
  t %>%
  kbl(.,
      format = "latex",
      caption = "Correlation Table and Descriptive Statistics",
      booktabs = TRUE,
      align = c("l", rep("c", ncol(.) - 1))) %>%
  kable_classic() %>%
  add_header_above(
    .,
    c(
      "",
      "Correlations" = ncol(allMlCor),
      "Descriptives" = nrow(allMlCor) - ncol(allMlCor)
    )
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person correlations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  kable_styling(latex_options = "scale_down") %>%
  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
  save_kable("Tables/descrWide.tex")

```

\subsection{Analysis and Results}

\subsubsection{Feature extraction}

\subsubsection{Feature reduction}

\subsubsection{Feature clustering}

feature engineering: feature extraction + feature selection

use domain knowledge to extract new variables from raw data (summarize)

options: tsfresh \citep[][]{christ2018}

\textit{k} features means $2^k – 1$ possible models

maximize relevance and minimize redundancy

```{r, include=FALSE}

```