### 2023-02-17 latex-results/05FeatureClustering.tex

For our illustration of the feature clustering procedure, we use the
generalizable and efficient centroid-based k-means clustering. The
procedure needs the user to make a selection on the optimal number of
clusters in order to run the main algorithm. We use the Elbow and the
Silhouette method, which both suggest two clusters to be the optimal
solution (see Supplemental Material A). We then entered the
participants' PC-scores into the k-means algorithm. We used the
recommended Hartigan and Wong algorithm \citep{hartigan1979} with 100
random initial centroid positions to avoid convergence to a sub-optimal
solution (i.e., local minima). The k-means analysis assigned 80
participants to cluster 1 and 76 participants to cluster 2.