### 2023-02-21 latex-results/02InputVariables.tex

For our illustration, we include 12 variables that were available across
all three studies and captured information about the participant's
interactions, as well as cognitive-, emotional-, and motivational self
in relationship with the majority group. We chose these aspects in
particular because (1) the interaction-specific information exemplified
the structural missingness issue of modern ESM data and (2) the
motivational, emotional, and cognitive experience offered a diverse
conceptualization of migration experience (beyond behavioral
measurements) that is becoming more common in the literature
\citep[][]{Kreienkamp2022d}. Full methodological details are available
in Online Supplemental Material A, but basic item information,
descriptives, and correlations are also available in
\tblref{tab:descrLong}.

Once the important variables have been selected, the data needs to be
prepared for the analysis steps. Importantly, this not only means
validating and cleaning the data (e.g., re-coding, removing duplicate or
unwanted measurements) but also making the time series comparable. Two
important steps are making the time-frames and response scales
comparable across participants --- for example, by choosing a time frame
that is common to most participants and standardizing the participants'
responses \citep{liao2005}.

In our illustration data set, the studies differed substantially in the
maximum length of participation (\(\text{max}(t_{S1})=\) 63,
\(\text{max}(t_{S2})=\) 69, \(\text{max}(t_{S3})=\) 155). This was
likely due to the option to continue participation without compensation
in the later two studies. To make the three studies comparable in
participation and time frames, we iteratively removed all measurement
occasions and participants that had more than 45\% missingness
\citep[which was in line with the general rcecommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
This procedure let to a final sample of 157 participants, who jointly
produced 8,132 measurements. Importantly, both the participant
repsonse-patterns and the time frame were now substantially more
comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
\(\text{max}(t_{S3})=\) 67). For a full overview of the data reduction
procedure and study-specific exclusions see Online Supplemental Material
A.

\input{tables/descrLong}