### 2023-02-23 latex-results/03FeatureExtraction.tex
:sparkles: feature extraction illustration choices added

\paragraph{Central tendency.}

The central tendency refers to the statistical measures that represent
the ``typical'' or ``average'' of a set of data. The most common
measures of central tendency are the mean, median, and mode
\citep{weisberg1992}. As a familiar statistic from probability theory,
the central tendency sits at the heart of many fundamental questions
about psychological time series. Researchers might, for example, be
interested in whether ``Over a one month period, are some people happier
than others?''. For our central tendency feature we chose the robust
\textit{median}, which can avoid potential issues with non-normally
distributed time series responses or outliers \citep{weisberg1992}. To
calculate the \textit{median} (\(M\)), we let \(X_{ij}\) be the ordered
list of values from the time series of variable \(j\) for participant
\(i\). The calculation depends on whether the number of measurements in
a time series \(n\) is odd or even.

\begin{equation} \label{eq:median}
  M(X_{ij}) = 
    \begin{cases}
      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
    \end{cases}
\end{equation}

\paragraph{Variability.}

Variability captures the degree to which a set of data differs from the
central tendency and is sometimes also referred to as the dispersion or
spread of the data \citep{weisberg1992}. In time series analyses,
variability is conceptually important because information about the
distribution and diversity of the data has been found to be indicative
of worse psychological states \citep{myin-germeys2018}. Person level
differences of ESM measurements have, for example, been associated with
higher levels of psycho-pathological recurrences among depression
patients \citep{timm2017}. For our illustration data, we chose to
capture the time series variability with the
\textit{Median Absolute Deviation} (\(MAD\)), where we calculate the
\textit{median} (\(M\); calculated as in \equatref{eq:median}) for the
absolute deviations of measurement \(x\) at time point \(t\) for
participant \(i\) and variable \(j\) from the median of that time series
\(X\).

\begin{equation} \label{eq:mad}
  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
\end{equation}

\paragraph{Instability.}

Instability captures the average change between two consecutive
measurements \citep{ebner-priemer2009}. While instability is
conceptually related to the variability feature, variability does not
take into account temporal dependency, whereas instability looks at the
`jumpy-ness' of the data over time. In other words, variability reflects
the range or diversity of values in a time series data, while
instability reflects the fluctuation or inconsistency in a time series
data over time \citep{trull2008}. For example, if a person has rapid and
extreme changes in mood their mood is highly unstable, while if a
person's mood responses span a wide range over the entire study period,
their mood is highly variable \citep{jahng2008}. Within psychological
time series, instability measurements have especially been important in
the research of borderline personality disorder \citep{trull2008} and
suicidality \citep{kivela2022}, but also in understanding early warning
signals more generally \citep{wichers2019}. Conceptually, the
instability feature, thus, relates to a broad range of research
questions, including: ``Do changes tend to be slow and gradual or fast
and abrupt?'' For our data we chose the \textit{mean absolute change}
\citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
the average absolute difference of measurements \(x\) at time points
\(t\) and \(t-1\), for each time series \(X\) of participant \(i\) and
variable \(j\).

\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t}-x_{t-1}\right|
\end{equation}

\paragraph{Self-similarity.}

Self-similarity in time series data refers to the property of a time
series to exhibit similar patterns of behavior over different time
scales \citep{dmello2021}. That is, self-similarity describes how much a
measurement carries over to future measurements. One important
self-similarity in psychological time series is \texit{inertia} --- how
much a measurement carries over to its next measurement
\citep{kuppens2010}. If inertia is high a development tends to stay in a
certain state or continues to move towards a certain direction. Because
high inertia is resistant to change, in emotion dynamics high inertia
has been found to be indicative of underreactive systems and to be
characteristic of psychological maladjustment \citep{kuppens2010}. In a
similar vein, high inertia at baseline was even predictive of the
initial onset of depression \citep{kuppens2012}. Conceptually, inertia
is more broadly connected to research questions such as: ``Do patients
stay in a depressed mood for several measurements?'\,' For our
illustration case, we chose the commonly used autocorrelation or
autoregression with a lag-1 to capture the inertia. High autocorrelation
values can indicate high levels of inertia, while low autocorrelation
values may indicate a more unpredictable or volatile time series
\citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
at the average correlation between a measurement \(x\) and the
proceeding measurement \(x_{t-1}\) for the time series \(X\) of
participant \(i\) and variable \(j\).

\begin{equation} \label{eq:ar}
  r_{ij,1} = \frac{\sum_{t=1}^{n-l}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
\end{equation}

\paragraph{Linear trend.}

In non-stationary time series a linear trend can be observed when there
is a consistent increase or decrease in the data over time
\citep{nyblom1986}. For psychological time series, researchers have, for
example, pointed out the importance of linear trends in interpersonal
communications \citep{vasileiadou2014}, and emotion dynamics
\citep{oravecz2016}. Theoretically, linear trends are often considered
the simplest way of assessing whether a psychological theory of change
is appropriate \citep{gottman1969}. In empirical practice, linear trends
are, thus, commonly exemplified by research questions such as ``Do
patient symptoms improve consistently?'' or ``Does worker productivity
decline continuously?'' For the variables in our illustration data set
we chose an overall linear regression slope to capture the linear trend.

\begin{equation} \label{eq:lin}
  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
\end{equation}

\paragraph{Nonlinearity.}

Changes in psychology are not always linear, instead nonlinearity is a
common feature of psychological time series \citep{hayes2007}. As an
example, episodic disorders, such as depression, are most likely best
described as non-linear systems \citep{hosenfeld2015}. Similarly,
patients in recovery from depression showed sudden changes in the
improvement of depression \citep{helmich2020a}. But also substance abuse
\citep{boker1998} or attitude changes rarely develop linearly
\citep{vandermaas2003}. Conceptually, researchers might have research
question about the type of the development: ``Is the development of
anxiety a nonlinear process?'' as well as the shape and structure of the
development: ``Does anxiety change as a sudden or smooth transition?''
We summarized the nonlinear trend with the
\textit{estimated degrees of freedom} of an empty GAM spline model. The
\(edf\) summarizes the \textit{wiggliness} of a spline trend line by
estimating the number of parametric and non-parametric (or smooth) terms
\(p\) in the model. An edf of 1 would be equivalent to a linear
relationship (i.e., one linear slope parameter), whereas a higher edf
(particularly an edf \textgreater{} 2) is indicative of a non-linear
trend.

\begin{equation} \label{eq:edf}
  edf \approx \sum p
\end{equation}

Beyond our main features of interest, we also extracted the
participant's number of completed ESM measurements to ensure that the
clusters are comparable in that regard. After the feature extraction, we
found that about 1.40\% of the extracted features are missing across the
72 features per participant. This might, for example, happen if
participants do not have two subsequent measurements with outgroup
interactions, so that an autocorrelation with lag-1 cannot be calculated
for the contact-specific variables. The small number of missing values
indicates that the feature-based approach indeed largely avoids the
structural missingness issue. The few missing values can, however, be an
issue for some feature reduction or feature clustering algorithms. We,
thus, impute the missing feature values with a single predictive mean
matching imputation using the MICE library.