### 2023-06-29 latex-results/07ClusterEvaluationInterpretation.tex
feedback LFB

We first inspect the clusters based on the average values of meaningful
features (see \fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We
see that for some variables the features are generally stronger in
separating the clusters. We, for example see that the item on
`\textit{how cooperative the interaction was}' distinguishes the two
clusters across almost all seven features (except for the
\textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}). Compare
this to the `\textit{outgroup attitudes}' item where the differences
between the clusters are much smaller for almost all features. We then
inspect the clusters with a focus on the features (see
\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
variable focus, we can see more clearly that some features are better at
distinguishing the clusters across variables. For example, \textit{MAD}
and \textit{median} distinguish the two clusters across almost all
variables (except for the item of whether the interaction was
representative of the outgroup). These two features stand in stark
contrast to other features, such as the
\textit{lag-1 auto correlations}, which showed much smaller differences
between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This
offers some information on which features were most important in
understanding the two extracted groups. Taken these two perspectives
together, we can also focus on individual features or variables in
particular. We, for example, see a strong difference in the average
well-being, where participants in cluster 2 showed a much lower median
well-being over the time series. At the same time, in terms of
stability, both groups have virtually identical averages \textit{MAC}
statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
are, thus, variables and features that distinguish the clusters better
than others and a combination of variables and features lets us explore
meaningful group differences in more detail. In our case, we see that
the central tendency, variability, and linear trend are best at
distinguishing a group with mainly positive experiences (cluster 1) from
a group with a more negative experience (cluster 2). We also see that
our clusters line up with the past literature on the importance of
focusing on simpler and more meaningful statistics
\citep{bringmann2018c, eronen2021a}.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons based on Features and Variables}
  \label{fig:clusterFeatVar}
  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
  \caption*{Note: \\
  Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
\end{figure}

In the second step, we look at prototypical trajectories of the
clusters. For k-means clustering it is often recommended to use the
average over time of the responses within the cluster
\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}. Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}. For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
Immediately striking are the mean differences, where participants in
cluster 1 had more meaningful and fulfilling outgroup interactions also
consistently reported more voluntary and cooperative interactions but
less accidental and involuntary interactions. The same cluster (cluster
1) also reported an increase in need-fulfilling interaction over the
30-day period and an increase in interactions that were representative
of the outgroup. Whereas the other cluster (cluster 2) showed a decrease
in voluntary, cooperative, and positive interactions over the 30 days.
This `deterioration' cluster (cluster 2) also saw a decrease in general
need fulfillment and experienced well-being over the 30 days (see
\fgrref[B]{fig:clusterTs}). We also see that while interaction
representativeness, outgroup attitudes, and well-being are relatively
stable for both clusters, the deteriorating cluster (cluster 2) also
showed substantially higher variablity and instability on most of the
other variables (also see \fgrref[A]{fig:clusterTs}).

Finally, we can also assess the clusters across other person-level
variable \citep[e.g.,][]{monden2022}. This out-of-feature comparison
allows us to check for data artifacts, as well as check whether the
developmental clusters are associated with important social markers and
individual differences. To illustrate artifact checks, we added the
number of ESM measurements into the comparison and find that the
participants in the deterioration cluster (cluster 2) on average
completed slightly more ESM surveys in general and reported on more
intergroup interactions in particular (see \(n\) in
\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures we
ensured that the general time frame and completion rates are similar for
all participants and indeed the numbers in ESM measurements generally
are largely similar (e.g., see \(n\) for well-being and outgroup
attitudes). However, the difference in reported number of interactions
might indicate either a clustering artifact or a meaningful difference.
The higher average number of interactions in cluster 2 could, for
example, indicate a clustering artifact if variances are more inflated
due to the larger samples \citep{kogan2006}. At the same time, however,
the difference in the number of experienced interactions might also
indicate a meaningful difference, where the deteriorating cluster
(cluster 2) on average reported more outgroup interactions, but these
interactions were less voluntary, less meaningful, and less positive.
Such a finding is in line with past research highlighting the role of
negative intergroup interations in explaining intergroup relations
\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. To further illustrate
the utility of assessing out of feature individual differences, we also
compare the two samples in terms of the participants' self-reported
discrimination experiences in the Netherlands (measured during the
post-measurement). When looking at the group comparison, we find that
participants in the deteriorating cluster (cluster 2) reported
substantially higher levels of everyday discrimination
(\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
(e.g., sum of specific ESM measurements) and cross-sectional variables
(e.g., general discrimination differences) that were not included in the
original clustering step can be used to explore and understand the
cluster differences in more detail.

In short, we find that that the feature-based clustering discerned two
meaningfully different groups of participants. We find an adaptive group
(cluster 2) that reports higher well-being and more positive outgroup
attitudes (\textit{median}) that are also stable over time
(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
period (\textit{linear trend}). This group also reported consistently
more meaningful, need-fulfilling, and cooperative outgroup interactions
(\textit{median}). This group with overwhelmingly positive experiences
stands in contrast with a more detrimental group (cluster 1). This
cluster, on average, reported much less positive, less meaningful, and
less fulfilling interactions and interaction patterns (\textit{median}).
This group also reported less positive outgroup attitudes, lower
well-being, and more discrimination experiences (\textit{median}). At
the same time, for members of this detrimental cluster conditions seemed
to deteriorate over time (\textit{linear trend}), and there was
generally less consistency in the experiences they were able to have
(\textit{MAC}, \textit{MAD}, \textit{edf}).

This cluster separation, then, has a number of empirical and practical
applications. Firstly, the clusters are descriptive. With tens of
variables, hundreds of participants, and thousands of measurements,
singular descriptive statistics are often not able to capture the
complex patterns that describe the data set. The feature-based
clustering offers some direct insight into the complexity within the
data set. In our empirical example, we for example see that participants
are meaningfully distinguished by a combination of high (vs.~low)
central tendency, variability, and linear trend. Secondly, the clusters
identify important groups. The adaptive and deteriorating groups offer
starting points for empirical exploration as well as practical
interventions. Researchers can start probing what exactly distinguishes
the two groups further and generate new bottom-up hypotheses.
Practitioners, can use the group separation to identify individuals in
need of assistance and can explore contextual factors that might
contribute to the difficulties some might face. Thirdly, the
feature-based approach is flexible and meaningful. We were able to use a
wide range of time series features that have been central in the ESM
literature and were able to use them directly to identify meaningful
groups. For our empirical illustration we, among others for example,
chose to focus on whether participants differed in their their average
well-being (i.e., \textit{median}), how much their well-being would vary
over time (i.e, \textit{MAD}), and whether their well-being would on
average increase or decrease over time (i.e., \textit{linear trend}).
Alternatively, for others cyclical patterns might be more important ---
for example, whether well-being was higher on weekends. Importantly, in
any case, we did not need to translate these dynamic features into
probabilistic inference models (e.g., VAR models) to cluster the
participants.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons over time}
  \label{fig:clusterTs}
  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
  \caption*{Note: \\
  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions.}
\end{figure}