### 2023-07-28 latex-results/01DataDescription.tex
feedback lfb

Matching these requirements, the data set we use consists of three
studies that followed migrants who had recently arrived in the
Netherlands in their daily interactions with the Dutch majority group
members \citep[for the data set see][]{Kreienkamp2022b}. After a general
migration-focused pre-questionnaire, participants were invited twice per
day to report on their (potential) interactions with majority group
members for at least 30 days. The short ESM surveys were sent out at
around lunch (12pm) and dinner time (7 pm). After the 30-day study
period participants filled in a post-questionnaire that mirrored the
pre-questionnaire. Participants received either monetary compensation or
partial course credits based on the number of surveys they completed.

The original studies included 207 participants (\(N_{S1}=\) 23,
\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 ESM
measurements. Each of the studies focused on recently arrived
first-generation migrants and each study included a number of
idiosyncratic variables relevant for the broader research collective.
For our empirical example, we focus on the variables that were collected
during the ESM surveys and were available in all three studies. Variable
selection and preparation are described as part of the illustration
below but for additional methodological details about the study setup
see \citet[][]{Kreienkamp2022b}.