### 2023-07-29 tables/var_selection.tex

\begin{sidewaystable*}[!hbtp]
\centering

\caption{\label{tab:var_selection}Variable Selection}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllccc}
\toprule
\multicolumn{4}{c}{ } & \multicolumn{2}{c}{Contact Specific} \\
\cmidrule(l{3pt}r{3pt}){5-6}
Variable & Question & Aspect & ESM & unspecific & Interaction Only\\
\midrule
AttitudesPartner & At the moment, how favorably do you feel towards -NAME- & cognition & \Checkmark &  & \Checkmark\\
InteractionContextAccidental & The interaction with -NAME- was accidental. & cognition & \Checkmark &  & \Checkmark\\
InteractionContextCooperative & The interaction with -NAME- was cooperative. & cognition & \Checkmark &  & \Checkmark\\
InteractionContextRepresentativeNL & The interaction with -NAME- was representative of the Dutch. & cognition & \Checkmark &  & \Checkmark\\
InteractionContextvoluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark &  & \Checkmark\\
\addlinespace
KeyNeedDueToPartner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark &  & \Checkmark\\
qualityMeaning & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & \Checkmark &  & \Checkmark\\
qualityOverall & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & \Checkmark &  & \Checkmark\\
AttitudesDutch & At the moment, how favorably do you feel towards the Dutch. & cognition & \Checkmark & \Checkmark & \\
DaytimeNeedFulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
\addlinespace
KeyNeedFulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark & \Checkmark & \\
exWB & How do you feel right now? very sad --- very happy & emotion & \Checkmark & \Checkmark & \\
\bottomrule
\end{tabular}}
\end{sidewaystable*}