### 2023-08-02 tables/var_selection.tex

\begin{sidewaystable*}[!hbtp]
\centering

\caption{\label{tab:var_selection}Variable Selection}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllccc}
\toprule
\multicolumn{4}{c}{ } & \multicolumn{2}{c}{Contact Specific} \\
\cmidrule(l{3pt}r{3pt}){5-6}
Variable & Question & Aspect & ESM & Interaction Only & unspecific\\
\midrule
Int: Accidental & The interaction with -NAME- was accidental. & cognition & \Checkmark & \Checkmark & \\
Int: Cooperative & The interaction with -NAME- was cooperative. & cognition & \Checkmark & \Checkmark & \\
Int: Meaningful & Overall, the interaction with -NAME- was: Superficial --- Meaningful & cognition & \Checkmark & \Checkmark & \\
Int: Need Fulfillment Partner & -NAME- helped fulfill your goal (-GOAL-) & needs & \Checkmark & \Checkmark & \\
Int: Partner Attitude & At the moment, how favorably do you feel towards -NAME- & cognition & \Checkmark & \Checkmark & \\
Int: Quality & Overall, the interaction with -NAME- was: Unplesant --- Plesant & cognition & \Checkmark & \Checkmark & \\
Int: Representative & The interaction with -NAME- was representative of the Dutch. & cognition & \Checkmark & \Checkmark & \\
Int: Voluntary & The interaction with -NAME- was voluntary. & cognition & \Checkmark & \Checkmark & \\
Int: Need Fulfillment & During your interaction with -NAME- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\
Need Fulfillment & During this -morning/afternoon- your goal (-GOAL-) was fulfilled. & needs & \Checkmark &  & \Checkmark\\
Outgroup Attitude & At the moment, how favorably do you feel towards the Dutch. & cognition & \Checkmark &  & \Checkmark\\
Well-Being & How do you feel right now? very sad --- very happy & emotion & \Checkmark &  & \Checkmark\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{6}{l}{\rule{0pt}{1em}All items used a continuous slider and were rescaled to a range of 0--100.}\\
\end{tabular}}
\end{sidewaystable*}