### 2023-08-07 supplemental-material-a-tutorial.qmd

---
title: "Tutorial: Feature-base time series clustering"
format:
  html:
    theme: cosmo
    # css: style.css
    toc: true
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
#| echo: true

1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).