### 2023-08-07 tutorial/about.qmd
tutorial about updated

---
title: "About"
bibliography: [packages.bib, ../referencesZotero.bib]
csl: ../apa.csl
---

```{r}
#| label: setup
#| include: FALSE

lib <- c(
  "rmarkdown",
  "knitr",
  "tidyverse",
  "kableExtra",
  "tsFeatureExtracR",
  "factoextra",
  "fpc",
  "readxl",
  "reshape",
  "tsFeatureExtracR",
  "stats",
  "plotly",
  "htmltools",
  "NbClust",
  "devtools",
  "pander",
  "knitcitations",
  "grateful",
  "bibtex"
)
invisible(lapply(lib, library, character.only = TRUE))
knitr::write_bib(file = 'packages.bib')

knitr::opts_chunk$set(
  class.source = 'numberLines lineAnchors'
)

```

This illustration accompanies the publication "A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series". The aim of this illustration website is to offer a transparent, open, and accessible repository providing annotated code for all analysis steps.

## Open Data

We provide all data necessary to replicate the full results of the main manuscript as well as this illustration. We specifically provide a minial reproducible dataset that includes all variables necessary to perform the analysis, while still protecting the privacy and confidentiality of the participants. All data is available as part of [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf].

## Open Code

The full code of this illustration as well as the main manuscript is available as part of [our GitHub repository](https://github.com) [@KreienkampMasked-clust-github]. A minimal version of the code as well as an archive file are available in [our Open Science ](https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4) [@KreienkampMasked-clust-osf]. For convenience purposes, we have also collected a number of functions we created as part of the illustration as an R-package. This package is available as its own GitHub repository and can be downloaded with the `devtools` package as `tsFeatureExtracR`: `devtools::install_github("JannisCodes/tsFeatureExtracR")` [@KreienkampBlinded2023-tsFeatureExtracR].

## Software Information

The full session information with all relevant system information and all loaded and installed packages is available in the collapsible section below.

<details>

<summary>

System Info

</summary>

\renewcommand{\arraystretch}{0.8}

<!-- decrease line spacing for the table -->

```{r Reproducibility-sessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
df_session_platform <- devtools::session_info()$platform %>%
  unlist(.) %>%
  as.data.frame(.) %>%
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kbl(
  df_session_platform,
  booktabs = T,
  align = "l",
  caption = "R environment session info for reproducibility of results" # complete caption for main document
)
```

\renewcommand{\arraystretch}{1}

<!-- reset row height/line spacing -->

</details>

<details>

<summary>

Package Info

</summary>

\renewcommand{\arraystretch}{0.6}

<!-- decrease line spacing for the table -->

```{r Reproducibility-sessionInfo-R-packages, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
df_session_packages <- devtools::session_info()$packages %>%
  as.data.frame(.) %>%
  filter(attached == TRUE) %>%
  dplyr::select(loadedversion, date, source) %>%
  rownames_to_column()

colnames(df_session_packages) <- c("Package", "Loaded version", "Date", "Source")

kbl(
  df_session_packages,
  booktabs = T,
  align = "l",
  caption = "Package info for reproducibility of results" # complete caption for main document
)
```

\renewcommand{\arraystretch}{1}

<!-- reset row height/line spacing -->

</details>

<details>

<summary>

Full Session Info (including loaded but unattached packages --- for troubleshooting only)

</summary>

`r pander(sessionInfo(), compact = FALSE)`

</details>

## References

```{r}
#| echo: false
#| results: asis

bib_entries <- read.bib("packages.bib")
all_keys <- names(bib_entries)

# Create a hidden div with all the references
cat('<div style="display:none;">')
for (key in all_keys) {
  cat(sprintf('[@%s] ', key))
}
cat('</div>')
```

::: {#refs}
:::