### 2023-08-11 TitlePage.tex

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper, donotrepeattitle]{apa7}

% change margins (e.g., for margin comments):
%\usepackage{geometry}
% \geometry{
% a4paper,
% marginparwidth=30mm,
% right=50mm,
%}

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}

% formatting links in the PDF file
\definecolor{googleurlcolor}{HTML}{1a0dab}
\hypersetup{
    pdfpagemode={UseOutlines},
    bookmarksopen=true,
    bookmarksopenlevel=0,
    hypertexnames=false,
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = googleurlcolor, %blue,Colour for external hyperlinks
    linkcolor    = black, %blue, Colour of internal links
    citecolor   = black, % cyan, Colour of citations
    pdfstartview={FitV},
    unicode,
    breaklinks=true,
}

% language settings
\DeclareLanguageMapping{american}{american-apa}

% custom open science badge height
\newlength{\badgeheight}
\setlength{\badgeheight}{1em}

% add reference library file
\addbibresource{referencesZotero.bib}

% Title and header
\title{A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series\\}
\shorttitle{feature-based time series clustering}

% Authors
\author[*,1]{Jannis Kreienkamp}
\author[1]{Maximilian Agostini}
\author[2]{Rei Monden}
\author[1]{Kai Epstude}
\author[1,3]{Peter de Jonge}
\author[1,3]{Laura F. Bringmann}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}
\affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
\affil[3]{Interdisciplinary Center Psychopathology and Emotion Regulation}

\authornote{   
   \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}\\
   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
   \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{Kreienkamp-clust-osf}. Materials and software are also available as part of our project GitHub repository \citep{Kreienkamp-clust-github}. The full illustration code is also available as a tutorial-style website at \href{https://tsfeatureclustr.com}{www.tsFeatureClustR.com} \citep{Kreienkamp-illustration-github}.

Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: \href{mailto:j.kreienkamp@rug.nl}{j.kreienkamp@rug.nl}

\vspace{1em}
\textit{Open Science Practices:}
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements

}

\leftheader{Kreienkamp}

\setlength\parindent{1.27cm}

% Start of the main document:
\begin{document}

% add title information (incl. title page and abstract)
%\maketitle

\makeatletter
\vspace*{2em}
\begin{center}
\noindent \textbf\@title
\vspace{1em}
\noindent \@author
\end{center}
\vspace{2em}
\begin{center}%
    \textbf\acksname%
\end{center}%
\vspace{-1em}
\protect\raggedright
\setlength{\parindent}{0.5in}
\indent\par\@acks%

\printbibliography

\end{document}