### 2023-08-11 manuscript_unmasked.tex
unmasked version added

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper, floatsintext]{apa7}

% change margins (e.g., for margin comments):
%\usepackage{geometry}
% \geometry{
% a4paper,
% marginparwidth=30mm,
% right=50mm,
%}

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome5}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath, nccmath}
\usepackage{mdframed}
\usepackage{ntheorem}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}
\usepackage{pdflscape}
\usepackage{lscape}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{tabu}
\usepackage{bbding}

% make warning with red triangle
\newcommand\Warning[1][2ex]{%
  \renewcommand\stacktype{L}%
  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries !}}{#1}}%

% make question with red triangle
\newcommand\Question[1][2ex]{%
  \renewcommand\stacktype{L}%
  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries ?}}{#1}}%

% add definition sections
\theoremstyle{break}
\newtheorem{definition}{Definition}

% add hypothesis sections
\theoremstyle{plain}
\theoremseparator{:}
\newtheorem{hyp}{Hypothesis}

\newtheorem{subhyp}{Hypothesis}
   \renewcommand\thesubhyp{\thehyp\alph{subhyp}}

% add quote section
\usepackage{csquotes}

% framed box section
\usepackage{framed}
\emergencystretch=1em

% formatting links in the PDF file
\definecolor{googleurlcolor}{HTML}{1a0dab}
\hypersetup{
pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = googleurlcolor, %blue,Colour for external hyperlinks
linkcolor    = black, %blue, Colour of internal links
citecolor   = black, % cyan, Colour of citations
pdfstartview={FitV},
unicode,
breaklinks=true,
}

% ref labels
\newcommand{\fgrref}[2][]{\hyperref[#2]{Figure \ref*{#2}#1}}
\newcommand{\tblref}[2][]{\hyperref[#2]{Table \ref*{#2}#1}}
\newcommand{\appref}[2][]{\hyperref[#2]{Appendix \ref*{#2}#1}}
\newcommand{\equatref}[2][]{\hyperref[#2]{Equation \ref*{#2}#1}}

% custom open science badge height
\newlength{\badgeheight}
\setlength{\badgeheight}{1em}

% prevent multipage footnotes
\interfootnotelinepenalty=10000

% language settings
\DeclareLanguageMapping{american}{american-apa}

% add reference library file
\addbibresource{referencesZotero.bib}

% Title and header
%\title{Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data}
\title{A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series}

\shorttitle{feature-based time series clustering}

% Authors
\author[*,1]{Jannis Kreienkamp}
\author[1]{Maximilian Agostini}
\author[2]{Rei Monden}
\author[1]{Kai Epstude}
\author[1,3]{Peter de Jonge}
\author[1,3]{Laura F. Bringmann}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}
\affil[2]{Hiroshima University, Graduate School of Advanced Science and Engineering}
\affil[3]{Interdisciplinary Center Psychopathology and Emotion Regulation}

\authornote{   
   \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
   \addORCIDlink{Rei Monden}{0000-0003-1744-5447}\\
   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}\\
   \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software are available at \url{https://janniscodes.github.io/migration-trajectories/} \citep{Kreienkamp-clust-github}. Materials, data, and code are available at \url{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4} \citep{Kreienkamp-clust-osf}.

Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: \href{mailto:j.kreienkamp@rug.nl}{j.kreienkamp@rug.nl}
}

\leftheader{Kreienkamp}

% Abstract
\abstract{
Psychological researchers and practitioners collect increasingly complex time series data aimed at identifying differences between the developments of participants or patients. Past research has proposed a number of `dynamic measures' that describe meaningful developmental patterns for psychological data (e.g., instability, inertia, linear trend). Yet, commonly used clustering approaches are often not able to include these meaningful measures (e.g., due to model assumptions). We propose feature-based time series clustering as a flexible, transparent, and well-grounded approach that clusters participants based on the dynamic measures directly using common clustering algorithms. We introduce the approach and illustrate the utility of the method with real-world empirical data that highlight common ESM challenges of multivariate conceptualizations, structural missingness, and nonlinear trends. We use the data to showcase the main steps of input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation. We also provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation.

}

\keywords{
    time series analysis, feature-based clustering, intensive longitudinal data, ESM\\
    \vspace{1em}
    \textit{Open Science Practices:}
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, \break
    \href{https://osf.io/75ryj/?view_only=db14031cd8704085b5c6165c9f5ed3e4}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
}

% set indentation size
\setlength\parindent{1.27cm}

% Start of the main document:
\begin{document}

% add title information (incl. title page and abstract)
\maketitle

% **CHEAT SHEET / LEGEND**

% Comments:
% '%' starts a comment in LaTeX (not printed)
% '\todo[inline]{} makes orange boxes in PDF
% '\marginpar{}' notes in margins
% '\footnote{}' footnote
% '\Warning' important note indicator in PDF (triangle with exclamation mark)
% '\Question' question note indicator in PDF (triangle with question mark)

% Citation (with Natbib citation style):
% '\citep[e.g.][p. 15]{CitationKey}' citation in parentheses "(e.g., Berry, 2003, p. 15)"
% '\citet{CitationKey}' citation in text "Berry (2003)"
% '\citealt' and '\citealp' alternate citation without parentheses
% '\citeauthor' and '\citeyear' only year or author

% Headings:
% '\part{}' and '\chapter{}' only relevant for multi-part or multi-chapter documents
% '\section{}' heading level 1
% '\subsection{}' heading level 2
% '\subsubsection{}' heading level 3
% '\paragraph{}' heading level 4
% '\subparagraph{}' heading level 5

% formatting:
% '\textbf{}' text bold font
% '\textit{}' text italic font
% '\underline{}' text underline
% '\sout{}' text strike out
% '\textsc{}' text small caps
% '\vspace{1em}' add vertical space
% '\hspace{1em}' add horizontal space
% '\\' new line (i.e., line break)
% '\pagebreak' start new page (i.e., page break)
% '\noindent' do not indent current line (e.g., current paragraph)
% 'begin{center}...end{center}' center text or object

% Math mode:
% '$\alpha = .8$' mathematical equation inline
% '$$\hat{y} = b_0 + b_1x$$' mathematical equation in its own line
% '\begin{equation}...\end{equation}' multi-line equation
% '\approx' approximate symbol
% '\neq' not equal
% '\bar' mean bar over letter
% '\pm' plus minus sign 
% '^{}' superscript
% '_{}' subscript
% '\fraq{numerator}{denominator}' fraction
% '\sqrt[n]{}' square root
% '\sum_{k=1}^n' sum for 1 through n

% Insert things from elsewhere:
% '\input{filename}' inputs the raw (tex) file as a command (e.g., tables and R-Markdown imports)
% '\include{filename}' includes section on new page (incl. possible auxiliary info)
% '\includegraphics[settings]{filename}' add a figure or graph
% '\caption{}' adds a caption to a table or figure
% '\label{}' labels sections, tables, figures, etc. so that they can be referred to.
% '\ref{}' refer to a labelled sections, tables, figures, etc.
% '\begin{enumerate}...\end{enumerate}' numbered list
% '\begin{itemize}...\end{itemize}' bullet-ed list
% '\item' item in list section

% Symbols:
% '\&' and sign
% '\%' percent sign
% '\_' three dotes
% '\#' hash symbol
% ------------------------------------------------------------------

\newlength{\mdfmar}
\setlength{\mdfmar}{1.5em}
\mdfdefinestyle{mdfbox}{
    innerleftmargin = +\mdfmar, %
    innerrightmargin = +2\mdfmar, 
    innertopmargin = +0\mdfmar, 
    innerbottommargin = +1.5\mdfmar, 
    skipabove = -12pt
}

% \begin{mdframed}[style=mdfbox]
% \noindent\center\textit{Structure}:
% \begin{itemize}[nosep]
%     \item introduction
%         \begin{itemize}[nosep]
%             \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
%             \item ESM literature: features to capture meaningful aspects
%             \item current clustering based on common model parameters, which only include a small selection of features and often have strong assumptions
%             \item Instead of building more complicated models (e.g., address violated assumptions or add different types of parameters), we propose to cluster based on important and relevant features directly.
%             \item Called feature-based clustering. Common procedure in digital phenotyping and the broader ml literature. Used in many fields
%             \item in this paper we provide a practical introduction to the method and illustrate its utility with real-world ESM data.
%         \end{itemize}
%     \item Data Used for Illustration
%     \item Analysis Steps + Application
%         \begin{itemize}[nosep]
%             \item input variables
%             \item feature extraction
%             \item (feature reduction)
%             \item feature clustering
%             \item cluster evaluation
%         \end{itemize}
%     \item Discussion (summary, limitation, implications, conclusion, etc.)
%     \newpage
%     \item\ [...]
%     \item Feature-based Time Series Clustering
%     \begin{itemize}[nosep]
%         \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
%         \item Features in Psychological Time Series
%         \begin{itemize}[nosep]
%             \item central tendency
%             \item variability
%             \item (in)stability
%             \item self-similarity
%             \item linear trend
%             \item nonlinearity
%         \end{itemize}
%         \item Analysis Steps
%         \begin{itemize}[nosep]
%             \item input variables
%             \item feature extraction
%             \item (feature reduction)
%             \item feature clustering
%             \item cluster evaluation
%         \end{itemize}
%     \end{itemize}
%     \item Empirical Illustration
%     \item Discussion (summary, limitation, implications, conclusion, etc.)
% \end{itemize}
% \vspace{1em}
% \end{mdframed}
% \newpage

Recent years have seen a striking increase in the number and variety of research studies that follow participants' everyday experiences and collect real-world psychological time series \citep[e.g.,][]{hamaker2017}. These intensive longitudinal data sets come with different sources of heterogeneity, where researchers have to consider differences across large numbers of participants, time points, and variables \citep[e.g.,][]{cattell1966, wardenaar2013}. However, despite its complexity, researchers are often interested in precisely this complexity and wish to understand how people differ in their developments across several variables \citep[e.g.,][]{ernst2021}. Researchers and practitioners are, for example, asking: ``Do the symptoms of different patients develop in contrasting ways?'' \citep{Monden2015} or ``How do migrants differ in the development of their self-reported needs as they arrive in a new country?'' \citep{Kreienkamp2022d}. There is, thus, a clear need for analysis techniques that identify between-subject differences in developmental patterns for psychological data.

Recently, one promising way of identifying between-subject developmental patterns has been \textit{time series clustering} --- the idea of inductively grouping participants based on similarities of their time series \citep{ariens2020}. This analysis type essentially seeks to capture comparable within-person developments --- such as whether a variable remains stable over time, consistently increases, or exhibits cyclical patterns --- and then groups the persons based on these patterns  \citep[][]{liao2005}\footnote{It should be noted in some cases the time series do not need to be summarized and can be compared directly. Such analyses are however only possible with highly regular and controlled data, such as EEG data \citep{huang1985}, or when clustering variables within the person rather than identifying differences between persons \citep{haslbeck2022}. Multivariate data from intensive psychological survey studies, as we discuss here, are seldomly directly comparable across persons \citep[e.g., ][]{faloutsos1994}. We provide a broader embedding of the current methods for psychological time series in the discussion section.}. Time series clustering, thus, crucially depends on identifying meaningful summaries of the time series developments, which can be used to compare participants \citep[][]{Aghabozorgi2015}.

Fortunately, past conceptual and empirical works in the experience sampling (ESM) literature have collected a number of meaningful aspects of psychological time series\footnote{In psychology, intensive longitudinal data collection methods are often referred to as experience sampling method (ESM), ecological momentary assessment (EMA), or ambulatory assessment (AA) studies. While the terms come from different conceptual backgrounds, they share a focus on collecting data over an extended period of time to capture people's behaviors and experiences as they vary over time and in response to different situations and events. In this article, we will use the experience sampling (ESM) term as it has the strongest footing within the clustering literature.}. Such aspects can be indicative of adaptive and maladaptive developments within the individual, can identify crucial transitions, or more generally are helpful in understanding a psychological time series. Important aspects might include concerns over whether a symptom consistently stays at a certain level without much variability or whether some emotions develop together. For many of the most important developmental aspects, researchers have assembled numeric measures that capture these patterns. These summary statistics are often called ``dynamic measures'', ``principles of change'', or ``dynamic features'' of the psychological time series \citep{dejonckheere2019, kuppens2017, krone2018}. Most research groups working on these time series features have proposed an overlapping number anywhere between four and twelve key features relevant to psychologists \citep[][]{wang2006, dejonckheere2019}. Each of these time series features not only captures a distinct aspect of psychological time series but also holds conceptual value --- inertia, for example, describes a resistance to change that can be indicative of psychological maladjustment \citep{kuppens2010} or a higher within-person variability can signal an erratic state \citep{myin-germeys2018}.

Yet, despite this diversity of meaningful time series features in psychology, most clustering of ESM data has only focused on a small and restrictive selection of time series characteristics. Thus far, the most common approach has been to cluster participants based on person-specific model parameters --- notably intercepts and slopes from vector autoregression models \citep[VAR; e.g.,][]{ariens2020, bulteel2016, stefanovic2022}. While such model parameters have the advantage of being familiar to researchers in the field, they are often restricted to autocorrelations and cross-lagged partial correlations \citep[e.g.,][]{bringmann2018c} --- only two of the many potentially important time series features. Additionally, the validity of model parameters is traditionally restricted by the assumptions of the particular statistical model. To take the common VAR model as an example, the model explicitly assumes that the time intervals between measurements are consistent (i.e., equidistant measurement assumption), it does not allow for missing values, and it assumes that means or variances do not change over time \citep[i.e., stationarity assumption;][]{lutkepohl2005}. These assumptions however stand in contrast to types of data researchers commonly collect to address important questions of erratic, and context-specific phenomena \citep[][]{myin-germeys2018, hamaker2017, kivela2022, helmich2020a}. Model parameters might, thus, not always accurately capture the time series and are often restricted in the time series features they capture.

% There is, thus, a resolute need for more flexible and meaningful time series summaries that can capture diverse measurement and missingness patterns and can capture non-stationary data, including non-linear trends, also across different time scales.

%Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models aimed at addressing either accuracy or interpretability concerns. Accuracy concerns have particularly been addressed by building more advanced models that relax specific assumptions \citep[e.g.,][]{denteuling2021} or by switching to different types of models with fewer assumptions, including spline models \citep{axen2011}. To expand interpretability, researchers have proposed to model additional types of parameters, including variability or granularity parameters \citep[e.g., see][]{krone2018}. 
%\citet{krone2018} even built a parametric model to tentatively cluster study participants.

Recent efforts to address the shortcomings of model parameters for time series clustering have primarily proposed more complicated models, which either seek to relax specific assumptions \citep[e.g.,][]{denteuling2021} or include additional time series features \citep[e.g., see][]{krone2018, gates2017}. In this manuscript, we instead, propose to directly cluster based on important and relevant time series features. Adequately named \textit{feature-based time series clustering}, such an approach has been a common procedure in digital phenotyping \citep[][]{loftus2022} and the broader machine learning literature \citep[][]{maharaj2019}. As such, the analysis has been applied to a variety of data, including analyses of astronomical, meteorological, and aviation pathways, biological and medical developments, as well as energy and finance patterns \citep{Aghabozorgi2015}. We argue that for psychological time series data, feature-based clustering offers a flexible approach, fewer strict assumptions, easy and intuitive analysis, as well as meaningful psychological interpretability.

In the sections below we seek to provide a practical introduction to the method. To do so, we illustrate the utility of the method with real-world ESM data. We use this data to discuss which psychological time series features are well-suited for a clustering approach, introduce the individual analysis steps, and provide practical guidance on common algorithms and analysis code.

\section{Data Used for Illustration}

To illustrate the functioning and utility of feature-based time series clustering with psychological ESM data, we apply the clustering process to a recent set of studies that collected data on migration experiences. Researchers have recently started using ESM data to follow the daily interactions of migrants with the cultural majority groups. We have seen both new technological advances to capture such interactions \citep[e.g.,][]{Keil2020} as well as a rise in empirical studies that look at aspects such as the well-being of migrants \citep[e.g.,][]{hendriks2016} and the distress of intergroup contact \citep[e.g.,][]{doucerain2023}.

This type of research comes as a response to a long-standing theoretical tradition highlighting the dynamic and developmental nature of cultural adaptation \citep[e.g.,][]{berry1986}. Importantly, such developmental trajectories are often difficult and stressful \citep{Ward2016} --- so that some may have an adaptive trajectory while others face a more difficult and grievous trajectory \citep{kim2017}. Prominent reviews within the migration literature have, thus, called for more longitudinal \citep[e.g.,][]{Ward2019} and real-world data \citep[e.g.,][]{McKeown2017}. By the same token, there is a crucial need to distinguish (mal-)adaptive clusters within these developmental trajectories and to relate these clusters to individual differences and contextual variables to make them understandable \citep[e.g.,][]{choi2009}.

At the same time, conceptual works on cultural adaptation have highlighted the multidimensional perspective necessary to understand migration experiences. A recent scoping review has particularly highlighted that cultural adaptation is best understood as a joint process of motivational, affective, cognitive, and behavioral aspects \citep[e.g.,][]{Kreienkamp2022d}. The data on migration experiences are, thus, explicitly multidimensional, mirroring the increase of complex data within the ESM literature \citep[][]{wardenaar2013}.

In short, the data on migration experience we are using for this illustration, address a pressing societal issue of identifying and understanding diverging trajectories. And importantly for our illustration, the migration ESM research, also, exemplifies the real-world data issues that ESM data commonly face, including a multivariate conceptualization with event-specific missingness patterns (also see \appref[]{app:ChallengesAppendix} for an expanded discussion of the current challenges within ESM data that are addressed within the data).

Matching these requirements, the data set we use consists of three
studies that followed migrants who had recently arrived in the
Netherlands in their daily interactions with the Dutch majority group
members \citep[for the data set see][]{Kreienkamp2022b}. After a general
migration-focused pre-questionnaire, participants were invited twice per
day to report on their (potential) interactions with majority group
members for at least 30 days. The short ESM surveys were sent out at
around lunch (12pm) and dinner time (7 pm). After the 30-day study
period participants filled in a post-questionnaire that mirrored the
pre-questionnaire. Participants received either monetary compensation or
partial course credits based on the number of surveys they completed.

The original studies included 207 participants (\(N_{S1}=\) 23,
\(N_{S2}=\) 113, \(N_{S3}=\) 71) with a total of 10,297 ESM
measurements. Each of the studies focused on recently arrived
first-generation migrants and each study included a number of
idiosyncratic variables relevant for the broader research collective.
For our empirical example, we focus on the variables that were collected
during the ESM surveys and were available in all three studies. Variable
selection and preparation are described as part of the illustration
below but for additional methodological details about the study setup
see \citet[][]{Kreienkamp2022b}. Each study was approved by the ethics
board of the university of origin and all participants gave informed
consent.

\section{Analysis Steps and Application}

To introduce and illustrate the feature-based clustering analysis, we will follow the conceptual steps of the procedure in sequential order and discuss key issues for each step. To do so we will follow the common separation that has structured feature-based clustering into four main steps \citep{rasanen2009, wang2006}. (1) The selection and preparation of the input variables, (2) the extraction of the time series features that describe the time series, with an optional feature reduction step if there are too many data points for the clustering algorithms, (3) the actual clustering of the time series features, and (4) the evaluation and interpretation of the clusters. 
While this is a common conceptual separation of procedural elements, it is important to note that these steps are a general outline, and the specific details of the analysis will depend on the nature of the data and the research question being addressed. Nonetheless, the conceptual nature of these steps allows us to introduce the major elements of the analysis. We also provide a conceptual overview that can be used alongside this section in \fgrref{fig:TSCFlow}.

\begin{figure}[!ht] % \begin{figure*}[hbtp] <-- on its own page
  \caption{Flowchart Feature-Based Time Series Clustering in Psychology}
  \label{fig:TSCFlow}
  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/TimeSeriesClusterFlowSelection.pdf}
  \caption*{Note: \\
  Choices selected for illustration in this manuscript are marked in bold.}
\end{figure}

\subsection{Input Variables}
Time series clustering starts with the selection and preparation of the variables of interest. While the selection will necessarily be field- and concept-specific, there are a few conceptual and methodological issues that should be considered. Conceptually, the included variables should adequately capture the concept of interest and should be meaningful to the understanding of the time series. One of the advantages of feature-based clustering is that it is inherently adept at accommodating multi-variate concepts --- a common aim in ESM research. There are, for example, calls that emotion dynamics should be assessed with a repertoire of positive and negative emotions \citep[e.g.,][]{dejonckheere2019}, many health developments are captured within the biopsychosocial domains \citep[e.g.,][]{suls2004}, and migration experiences are fully captured with affect, behavior, cognition, and desire measurements \citep[e.g.,][]{Kreienkamp2022d}. At the same time, however, the added number of variables can become a methodological concern. Not only can redundant and irrelevant variables diminish the quality of the analyses, but with intensive longitudinal data the number of data points compounds across participants, measurement occasions, and variables so that additional variables can make many of the following steps substantially more difficult (also see \fgrref{fig:TSCFlowN}).

\begin{figure}[!ht] %hbtp
  \caption{Exemplary Flowchart of Data Points in Feature-Based Time Series Clustering}
  \label{fig:TSCFlowN}
  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustFlowN.pdf}
  \caption*{Note: \\
  The presented number of participants, variables, and measurement occasions are somewhat arbitrary but generally represent common sample sizes found within the literature. Also, the number of extracted clusters is presented for illustrative purposes only.}
\end{figure}

For our illustration, we include 12 variables that were measured as part
of the ESM surveys in all three studies and captured information about
the participant's interactions, as well as cognitive-, emotional-, and
motivational self in relationship with the majority group (see
\tblref{tab:var_selection} for an overview). We chose these aspects in
particular because (1) the interaction-specific information exemplified
the structural missingness issue of modern ESM data and (2) the
motivational, emotional, and cognitive experience offered a diverse
conceptualization of migration experience (beyond behavioral
measurements) that is becoming more common in the literature
\citep[][]{Kreienkamp2022d}. The breadth of the included variables also
showcases the utility of the method for a growing body of literature
that considers heterogeneous and complex concepts. As a result, the
number of included variables is also on the higher end for psychological
concepts and additionally allows us to showcase the efficiency benefits
of the method and offers a reasonable use case for the optional feature
reduction step.

\input{tables/var_selection}

Once the important variables have been selected, the data needs to be
prepared for the analysis steps. Importantly, this not only means
validating and cleaning the data (e.g., re-coding, removing duplicate or
unwanted measurements) but also making the time-series comparable. Two
important steps are making the time-frames and response scales
comparable across participants --- for example, by choosing a time frame
that is common to most participants and standardizing the participants'
responses
\citep['data exclusion' and 'data transformation' in \fgrref{fig:TSCFlow}; also see][]{liao2005}.

In our illustration data set, the studies differed substantially in the
maximum length of participation (\(\text{max}(t_{S1})=\) 63,
\(\text{max}(t_{S2})=\) 69, \(\text{max}(t_{S3})=\) 155). This was
likely due to the option to continue participation without compensation
in the latter study. To make the three studies comparable in
participation and time frames, we iteratively removed all measurement
occasions and participants that had more than 45\% missingness
\citep[which was in line with the general recommendation for data that might still need to rely on imputations for later model testing][]{Madley-Dowd2019}.
This procedure led to a final sample of 157 participants, who jointly
produced 8,132 measurements. Importantly, both the participant
response-patterns and the time frame were now substantially more
comparable (\(\text{max}(t_{S1})=\) 61, \(\text{max}(t_{S2})=\) 60,
\(\text{max}(t_{S3})=\) 67). Full methodological details are available
in Online Supplemental Material A, but basic item information,
descriptives, and correlations are also available in
\tblref{tab:descrLong}.

\input{tables/descrLong}

\subsection{Feature Extraction}
Armed with a relevant selection of key variables, the main aim of the feature extraction is to describe the most important and meaningful aspects of a time series. In its most general approach, feature extraction can include any numeric summary of the time series \citep[e.g.,][]{maharaj2019}. Given this flexibility, a staggering variety of time series features have been proposed across different disciplines. For example, \citet{wang2006} proposed 9 structural characteristics \citep[also see][]{fulcher2013}, \citet{adya2001} collected 28 features relevant for forecasting, and a commonly used software package for feature extraction `tsfresh' allows users to extract a total of 794 features of a time series \citep[][]{christ2018}.

However, not all time series features might be relevant to psychological time series or any particular research question. For example, a psychologist interested in well-being might not necessarily be interested in the exact time point after which 50\% of the summed well-being values lie (i.e., relative mass quantile index) or how much different sine wave patterns within the well-being data correlate with one another (i.e., cross power spectral density). Instead, we advocate that we look at time series features that have a strong backing within the ESM literature and offer meaningful interpretability.

Fortunately, past conceptual and empirical efforts offer valuable discussions of common time series features in psychological research. To understand emotion dynamics, \citet{kuppens2017} originally proposed four dynamic features: (1) within-person variability, (2) co-variance or intraclass coefficient (ICC), (3) inertia or autocorrelation, and (4) cross-lagged correlations. These features were then extended by \citet{krone2018}, adding (5) innovation variance, and (6) mean intensity. \citet{krone2018} even built a parametric model to tentatively cluster study participants. From a slightly different perspective \citet{dejonckheere2019} later added three additional features for psychological time series: (7) instability (8) interdependence (i.e., network density), and (9) diversity \citep[i.e., Gini coefficient; also see][]{wendt2020}\footnote{It should be noted that also within the psychological literature, alternative summaries have been proposed that, for example, include measurement distribution, nonlinear developments, or categorical states. As an example, \citet{kiwuwa-muyingo2011} proposed to extract clinically meaningful states for medical adherence data and suggests these states as meaningful time series features.}.

Some of the time series features found in the psychological literature are not necessarily well-suited to summarize time series for feature-based clustering and some key conceptual features are not well represented in the literature. In particular, covariances and cross-lagged correlations often produce a large number of parameters and might not necessarily summarize the existing data enough \citep{ernst2021}. Others such as network density parameters, used to summarize variable interdependence, might not always be meaningful for psychological data \citep{bringmann2019a}. At the same time, linear and nonlinear trends are not captured by the features commonly proposed for psychological time series, because the features are often developed for stationary models \citep[e.g.,][]{krone2018}.

Thus, while the final selection of time series features should always be driven by the research questions and field-specific conventions, for our illustration we chose six time series features that relate to common psychological research questions and recent works within the field: (1) central tendency, (2) variability, (3) instability, (4) self-similarity, (5) linear trend, and (6) nonlinearity. An overview of the selected time series features, their substantive interpretations, and mathematical operationalizations is available in \tblref{tab:esmFeatures}. For each of the six time series features, we selected a mathematical representation that was appropriate for our type of data. We provide a short introduction of each feature below. Beyond the operationalizations we chose for our case study, we collected the R-functions we created for the analyses as an R package that automatically extracts and prepares a large selection of the time series feature operationalizations presented in \tblref{tab:esmFeatures}. All functions are available as part of the package GitHub repository (see the \texttt{featureExtractor()} function; \citealp{Kreienkamp2023-tsFeatureExtracR}) and are annotated as part of our tutorial-style illustration (see Supplemental Material A).

\input{tables/esmPsychFeatures}

\paragraph{Central tendency.}

The central tendency refers to the statistical measures that represent
the ``typical'' or ``average'' of a set of data. The most common
measures of central tendency are the mean, median, and mode
\citep{weisberg1992}. As a familiar statistic from probability theory,
the central tendency sits at the heart of many fundamental questions
about psychological time series. Researchers might, for example, be
interested in whether ``Over a one-month period, are some people happier
than others?''

For the central tendency feature of our illustration we chose the robust
\textit{median}, which can avoid potential issues with non-normally
distributed time series responses or outliers \citep{weisberg1992}. To
calculate the \textit{median} (\(M\)), we let \(X_{ij}\) be the ordered
list of values from the time series of variable \(j\) for participant
\(i\). The calculation depends on whether the number of measurements in
a time series \(n\) is odd or even.

\begin{equation} \label{eq:median}
  M(X_{ij}) = 
    \begin{cases}
      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
    \end{cases}
\end{equation}

\paragraph{Variability.}

Variability captures the degree to which a set of data differs from the
central tendency and is sometimes also referred to as the dispersion or
spread of the data \citep{weisberg1992}. In time series analyses,
variability is conceptually important because information about the
distribution and diversity of the data has been found to be indicative
of worse psychological states \citep{myin-germeys2018, helmich2021}.
Person-level differences of ESM measurements have, for example, been
associated with higher levels of psycho-pathological recurrences among
depression patients \citep{timm2017}. As such, psychological researchers
and practitioners are often empirically interested in between-person
differences in variability. Researchers on polarization and
radicalization might for example ask: ``Are people settled in their
attitudes towards migrants or do they vary across the measurement
period?''

For our illustration data, we chose to capture the time series
variability with the \textit{Median Absolute Deviation} (\(MAD\)), where
we calculate the \textit{median} (\(M\); calculated as in
\equatref{eq:median}) for the absolute deviations of measurement \(x\)
at time point \(t\) for participant \(i\) and variable \(j\) from the
median of that time series \(X\). We again chose the robust statistic
because the Median-based measure is less affected by non-normal
distributions and extreme values or outliers compared to other measures
of variability like the standard deviation \citep{weisberg1992}

\begin{equation} \label{eq:mad}
  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij}) \right|)
\end{equation}

\paragraph{Instability.}

Instability captures the average change between two consecutive
measurements \citep{ebner-priemer2009}. While instability is
conceptually related to the variability feature, variability does not
take into account temporal dependency, whereas instability looks at the
`jumpy-ness' of the data over time. In other words, variability reflects
the range or diversity of values in a time series data, while
instability reflects the fluctuation or inconsistency in a time series
data over time \citep{trull2008}. For example, if a person has rapid and
extreme changes in mood their mood is highly unstable, while if a
person's mood responses span a wide range over the entire study period,
their mood is highly variable \citep{jahng2008}. Within psychological
time series, instability measurements have especially been important in
the research of borderline personality disorder \citep{trull2008} and
suicidality \citep{kivela2022}, but also in understanding early warning
signals more generally \citep{wichers2019}. Conceptually, the
instability feature, thus, relates to a broad range of research
questions, including: ``What is the nature of the identification changes
in those who start working in a new country?'' or ``Do strong daily
fluctuations in self-esteem reflect the process of identity formation in
adolescents?''

For our data we chose the \textit{mean absolute change}
\citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
the average absolute difference of two consecutive measurements \(x\) at
time points \(t\) and \(t-1\), for each time series \(X\) of participant
\(i\) and variable \(j\).

\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=2, \ldots, t}\left|x_{t}-x_{t-1}\right|
\end{equation}

Another common measurement of instability is the
\textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is
often preferred where differences in magnitude are more important than
the frequency of those changes, for example, when big shifts in time
series are considered more impactful or when outliers are meaningful and
need to be taken into account \citep{chatfield2003}.

\paragraph{Self-similarity.}

Self-similarity in time series data refers to the property of a time
series to exhibit similar patterns of behavior over different time
scales \citep{dmello2021}. That is, self-similarity describes how much a
measurement carries over to future measurements. One important
self-similarity in psychological time series is \textit{inertia} --- how
much a measurement carries over to its next measurement
\citep{kuppens2010, suls1998}. If inertia is high a development tends to
stay in a certain state. Because high inertia is resistant to change, in
emotion dynamics high inertia of negative affect has been found to be
indicative of under-reactive systems and to be characteristic of
psychological maladjustment \citep{kuppens2010}. In a similar vein, high
inertia in negative affect at baseline was even predictive of the
initial onset of depression \citep{kuppens2012}. Conceptually, inertia
is more broadly connected to research questions such as:
\texttt{Do\ patients\ stay\ in\ a\ negative\ mood\ for\ several\ measurements?\textquotesingle{}\textquotesingle{}\ or}Do
migrants stay with their language practice for several days at a
time?'\,'

For our illustration case, we chose the commonly used autocorrelation or
autoregression with a lag-1 to capture the inertia. High autocorrelation
values can indicate high levels of inertia, while low autocorrelation
values may indicate a more unpredictable or volatile time series
\citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
at the average correlation between a measurement \(x\) and the preceding
measurement \(x_{t-1}\) for the time series \(X\) of participant \(i\)
and variable \(j\) with \(n\) measurements.

\begin{equation} \label{eq:ar}
  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
\end{equation}

Where \(\overline{x}_{ij}\) is the mean of the time series \(x_{ij}\),
calculated as:

\begin{equation} \label{eq:mean_for_ar1}
  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
\end{equation}

\paragraph{Linear trend.}

In non-stationary time series, a linear trend can be observed when there
is a consistent increase or decrease in the data over time
\citep{nyblom1986}. For psychological time series, researchers have, for
example, pointed out the importance of linear trends in interpersonal
communications \citep{vasileiadou2014}, and emotion dynamics
\citep{oravecz2016}. Theoretically, linear trends are often considered
the simplest way of assessing whether a psychological theory of change
is appropriate \citep{gottman1969}. In empirical practice, linear trends
are, thus, commonly exemplified by research questions such as ``Do
patient symptoms improve consistently?'' or ``Does worker productivity
decline continuously?''

For the variables in our illustration data set, we chose an overall
linear regression slope to capture the linear trend. The regression
slope \(b_{ij}\) provides the average change from one time point \(t\)
to the next across all measurements \(x\) of a time series \(X\) of
participant \(i\) and variable \(j\). The specific form of the OLS slope
formula we provide below calculates \(b_{ij}\) as the sum across all
time points of the product of the deviation of time \(t\) from its mean
\(\overline{t}\) and the deviation of \(x_{ij}\) from its mean
\(\overline{x}_{ij}\) at each time point, divided by the sum across all
time points of the square of the deviation of time from its mean
(\(\sum(t-\overline{t})^2\)). Intuitively, the formula captures the rate
of change of variable \(x_{ij}\) with respect to time. This slope will
indicate how the variable \(x_{ij}\) changes over time, controlling for
its mean value and the mean of time. If the slope is positive,
\(x_{ij}\) increases over time; if it's negative, \(x_{ij}\) decreases
over time.

\begin{equation} \label{eq:lin}
  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
\end{equation}

\paragraph{Nonlinearity.}

Changes in psychology are not always linear, instead, nonlinearity is a
common feature of psychological time series \citep{hayes2007}. As an
example, episodic disorders, such as depression, are most likely best
described as non-linear systems \citep{hosenfeld2015}. Similarly,
patients in recovery from depression showed sudden changes in the
improvement of depression \citep{helmich2020a}. But also substance abuse
\citep{boker1998} or attitude changes rarely develop linearly
\citep{vandermaas2003}. Conceptually, researchers might have research
questions about the type of the development: ``Is the development of
well-being a nonlinear process?'' as well as the shape and structure of
the development: ``How many spikes in well-being did a migrant
experience?''

We summarized the nonlinear trend with the
\textit{estimated degrees of freedom} of an empty GAM spline model. The
\(edf\) summarizes the \textit{wiggliness} of a spline trend line
\citep{wood2017}. The degrees of freedom of a spline model are primarily
determined by the number of knots and the order of the spline. For
instance, a cubic spline with \(k\) knots has \(k\)+3 degrees of freedom
\citep{faraway2016}. However, in a penalized spline framework, which is
commonly used for GAMs, the effective degrees of freedom can be less
than \(k\)+3. This is because the model employs a smoothing parameter to
control the trade-off between the complexity (flexibility) of the model
and its fit to the data, thereby penalizing overly complex models and
potentially reducing the effective degrees of freedom \citep{marx1998}.
Intuitively then an edf of 1 would be equivalent to a linear
relationship (i.e., one linear slope parameter), whereas a higher edf
(particularly an edf \textgreater{} 2) is indicative of a non-linear
trend. The estimated degrees of freedom are commonly based on a concept
called `effective degrees of freedom' and can be represented as the
trace \(tr\) (i.e., the sum of the diagonal elements) of the smoother
matrix \(S\), a symmetric matrix that maps from the raw data to the
smooth estimates \citep{wood2017}.

\begin{equation} \label{eq:df}
  edf = tr(S)
\end{equation}

Beyond our main features of interest, we also extracted the
participant's number of completed ESM measurements to ensure that the
clusters are comparable in that regard (i.e., to exclude spurious
explanations for the cluster assignments). After the feature extraction,
we found that about 1.40\% of the extracted features are missing across
the 72 features per participant. This might, for example, happen if
participants do not have two subsequent measurements with outgroup
interactions, so that an autocorrelation with lag-1 cannot be calculated
for the contact-specific variables. The small number of missing values
indicates that the feature-based approach indeed largely avoids the
structural missingness issue. However, even the few missing values can
be an issue for some feature reduction or feature clustering algorithms.
We, thus, impute the missing feature values with a single predictive
mean matching imputation using the MICE library \citep[][]{buuren2011}.
Note again that with this procedure we only need to impute an extremely
small number of missing values as most feature calculations can use the
available data instead.

It is important to reiterate that the six selected time-series features are in no way exhaustive or imperative. Both using a more data-driven approach to the selection of time series features or selecting entirely different aspects to summarize the time series are legitimate options \citep[e.g., see][]{heylen2016}. Our choice seeks to offer a practical toolbox of time series features that are common and meaningful to psychological research questions and practice but are also easy to extract and summarize a broad range of developments without asserting strict assumptions.

\subsection{Feature Reduction}
Once a meaningful set of time series features has been extracted for each variable and participant, the total number of data points sometimes remains too large for the desired clustering algorithm. As an example, a relatively common scenario would include 10 variables of interest, where eight time series features are extracted, resulting in 80 features per participant (with a common sample size of 100 participants that would result in a total of 8,000 data points in this hypothetical example). We offer an illustration of the compounding numbers of data points in \fgrref{fig:TSCFlowN}. The difficulty of finding stable clusters for data with a large number of dimensions is sometimes termed the `dimensionality curse' \citep[e.g.,][]{altman2018}.

To deal with this dimensionality issue, two main approaches have been proposed --- feature selection and feature projection \citep[e.g.,][]{erdogmus2008}. While feature selection refers to the process of identifying and selecting a subset of relevant features from the original feature set \citep{alelyani2014}, feature projection refers to the process of transforming the original feature set into a new feature set of lower dimensionality \citep[][]{carreira-perpinan1997}. In general, feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample. Feature projection methods, on the other hand, have been popular because they are efficient, widely available, and applicable to a wide range of data types. We provide an overview of the common approaches, an intuitive introduction to common methods, and exemplar algorithms in \tblref{tab:featureReduction}.

%Feature selections seek to create a subset of the most important features. The many available approaches differ in how they seek to determine the importance of the individual features \citep{alelyani2014}. Generally speaking, selection methods can be categorized as `filter', `wrapper', or `hybrid' methods \citep{hoque2014}. The filter methods, broadly speaking determine important features by identifying irrelevant features (e.g., because features do not capture much information), and identifying redundant features \citep[e.g., because features capture the same information; e.g., see][]{yu2004}. Wrapper methods, on the other hand, avoid the feature-based evaluation and focus on the performance of the later model to identify the most important features \citep{kohavi1997}. A wrapper, thus, compares the performance of models with different feature combinations \citep[e.g., see][]{tang2014}. Traditional examples of wrappers are forward selection or backward elimination procedures. Because filter methods might not always perform well and wrapper methods are computationally intensive\footnote{Computationally \textit{k} features could be considered in \(2^k – 1\) possible combinations --- for the example of 80 features that would allow for \(1.20 \times 10^{24}\) (over one septillion) combinations.}, hybrid methods seek to combine the two methods and find a balance between computational effort (i.e., efficiency) and performance \citep[i.e., effectiveness; e.g.,][]{alelyani2014}. Methods might, for example, use a filter step to reduce the size of features considered in a later wrapper step \citep{hsu2011}. Feature selection procedures have the benefit that they retain the interpretable feature labels directly and immediately indicate which features were most informative in the sample.

%The second general approach to the dimensionality curse has been feature projection. Broadly speaking, projection methods seek to transform the many features in such a way that a much smaller number of new variables can accurately capture the variance and structure of the original features \citep[i.e., the data is projected to a lower dimensional space; e.g.,][]{carreira-perpinan1997}. Projection methods are relatively common in psychological research --- including, factor- and principal component analyses \citep{vandermaaten2009}. Commonly, this is achieved using linear transformations \citep[e.g., principal component analysis; see][]{cunningham2015}, or more complex nonlinear transformations \citep[e.g., t-SNE; see]{lee2007}. Feature projection methods have been popular because of their widespread availability as well as their high efficiency.

\input{tables/FeatureReduction}

For our own illustration data, we chose a feature projection method to
reduce the dimensionality of our extracted features. We particularly
chose the feature projection method for its broad applicability. We,
specifically, selected the commonly used
\textit{principal component analysis} (PCA). Some of the more
tailor-made feature selection algorithms can be more accurate in
reducing the feature dimensionality and might retain feature importance
information more directly, depending on the specific data structure.
However, PCAs have the distinct benefit that they are well-established
within the psychometric literature \citep{jolliffe2011} and can broadly
be applied to a wide variety of studies in an automatized manner
\citep{abdi2010}. As our aim is to present a general illustration that
can also be adopted across use cases, we present the workflow using a
PCA here but we encourage users to consider more specialized methods as
well.

To use the PCA with our extracted time series features, we first
standardize all features across participants to ensure that all features
are weighted equally. We then enter all 72 features into the analysis.
The PCA uses linear transformations in such a way that the first
component captures the most possible variance of the original data
\citep[e.g., by finding a vector that maximizes the sum of squared distances][]{jolliffe2002, abdi2010}.
The following components will then use the same method to iteratively
explain the most of the remaining variance while also ensuring that the
components are linearly uncorrelated \citep{shlens2014}. In practice,
this meant that the PCA decomposed the 72 features into 72 principal
components but now (because of the uncorrelated linear transformations)
the first few principal components will capture a majority of the
variance. We can then decide how much information (i.e., variance) we
are willing to sacrifice for a reduced dimensionality. A common rule of
thumb is to use the principal components that jointly explain 70--90\%
of the original variance
\citep[i.e., cumulative percentage explained variance; e.g.,][]{jackson2003}.
For our illustration, we select the first 27 principal components that
explain 80\% of the variance in the original 72 features (reducing the
dimensionality by 62.50\%). For the extracted principal components we
save the 27 principal component scores for each participant (i.e., the
participants' coordinates in the reduced dimensional space; PC-scores).

We would like to comment on two practical matters when using principal
components --- the amount of dimensionality reduction and the
interpretation of the principal components. As for the expected
dimensionality reduction, given its methodology, PCAs tend to `work
better' at reducing dimensions with (highly) correlated variables
\citep[e.g.,][]{jolliffe2002}. Thus, with a set of very homogeneous
variables and features users will need fewer principal components to
explain a large amount of variance, while a more diverse set of
variables and features will tend to require more principal components to
capture the same amount of variance \citep[e.g.,][]{abdi2010}. Our 27
principal components are still a relatively high number of variables but
this is not surprising as we chose a diverse conceptualization and a
diverse set of time series features. As for interpretability, PCA allows
users to extract information on the meaning of the principal components.
In particular, because the principal components are linear combinations
of the original features, users can extract the relative importance of
each feature for the extracted principal components (i.e., the
eigenvectors). While this can be useful in understanding the variance in
the original data or help with manual feature selection, we use the PCA
purely to reduce the dimensionality for the clustering step. Instead of
relying on the principal components, we use the original features of
interest to interpret the later extracted clusters. We particularly
advocate for such an approach if all original features are considered
meaningful in understanding the time series and users would like to
retain the features for interpretation (irrespective of the features'
importance).

\subsection{Feature Clustering}
For the actual clustering of the time-series features, the main aim is to organize participants into groups so that the features of participants within a group are as similar as possible, while the features of people in different groups are as different as possible \citep{liao2005}. The crux of clustering is, thus, to have clearly defined and effective measurements of (dis)similarity. Most of the clustering algorithms used today use some form of distance measurement to optimize group assignment \citep[or similarity measurement for qualitative features; see][]{Aghabozorgi2015}. While others have produced excellent overviews of the many clustering approaches available \citep[e.g.,][]{xu2015}, the more readily available approaches suitable for most time series feature data can, broadly speaking, be categorized as based on (1) centroids, (2) distributions, (3) density, (4) hierarchies, or (5) a combination thereof \citep[see \tblref{tab:clusterApproaches} for an overview; also see][for a broader review]{jain1999}.

There is, unfortunately, no one-size-fits-all solution to clustering and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency. We provide a short intuitive explanation for common approaches, together with some of their characteristics and example algorithms in \tblref{tab:clusterApproaches}. For our own illustration, we have chosen the centroid-based k-means clustering. Although k-means sacrifices some level of accuracy, it offers certain advantages. We specifically chose k-means because it is an extremely efficient method that works well with large participant- and feature numbers without making too many restrictive assumptions about the shape of the clusters \citep{jain2010}. K-means is also well-established within the research community and has been readily implemented in many statistical software packages \citep{hand2005}. Additionally, many of the feature selection methods have specifically been designed for the well-established k-means algorithm \citep[e.g.,][]{boutsidis2010}. As such, the k-means offers a good starting point for many psychological researchers and the method should be generalizable across a relatively wide variety of projects.

%Each of these approaches can be a valuable clustering approach for time series feature data and users will usually have to make an informed decision based on the structure of their data as well as an appropriate weighing of accuracy and efficiency. As an example, under ideal conditions, most approaches are likely to provide a similar cluster solution \citep[e.g., for well-separated groups with little noise and few outliers; e.g., see][]{peng2022}. However, when the shapes of clusters, for example, become more complex in a multi-dimensional space, density-based or hierarchy-based approaches that allow for more bottom-up clustering are likely to be more accurate \citep[e.g.,][]{langfelder2008}. Yet, with higher numbers of participants and features, both density- and hierarchy-based approaches may perform less well and the more efficient centroid-based methods might be more effective \citep[e.g.,][]{jain2010}. Similarly, several of the proposed time series features are statics that are generated from Gaussian distributions and a distribution-based algorithm might be ideally suited to separate such distributions \citep[e.g.,][]{corduas2008}. Yet, in cases where misspecifications are costly, a method with fewer assumptions might be more advisable \citep[e.g.,][]{ankerst1999}.

\input{tables/clusterApproaches}

During the k-means clustering itself, the analysis seeks to minimize the
total within-cluster variation. The analysis is designed to optimize the
clustering of the feature data into \(k\) groups, where \(k\) is a
pre-defined number of clusters. We used the Hartigan and Wong algorithm,
which is a widely used algorithm in k-means clustering
\citep{hartigan1979}. The algorithm starts by randomly separating the
data points into k clusters and then iteratively updates the assignment
of each point to the nearest cluster center until convergence. To do so,
the Hartigan and Wong algorithm specifically calculates the
within-cluster variation (\(W\)) of cluster \(C_i\) as the summed
squared Euclidean distances of the feature \(x\) to the closest cluster
centroid \(\mu_i\):

\begin{equation} \label{eq:kWCi}
  W(C_i) = \sum_{x \in C_i}(x-\mu_i)^2
\end{equation}

By summing the within-cluster sum of squares from all \(k\) clusters, we
can then derive the total within-cluster sum of square \(WCSS\):

\begin{equation} \label{eq:kWCSS}
  WCSS = \sum_{i=1}^k W(C_i) = \sum_{i=1}^k \sum_{x \in C_i} (x - \mu_i)^2
\end{equation}

It is this \(WCSS\) that becomes the objective function to be minimized,
by iteratively moving features from one cluster to another
\citep{hartigan1979}. In particular, the algorithm (1) calculates the
cluster centroids of the initial partitioning, (2) checks whether any
feature has a centroid that is closer than that of the currently
assigned cluster (3) updates the centroids based on any reassigned
features, and then iterates between steps two and three until \(WCSS\)
is minimized (i.e., locally optimal convergence) or a maximum number of
iterations is reached \citep{jain2010}. Given the iterative nature of
the algorithm, the initial partitioning is often important because the
algorithm might arrive at a suboptimal clustering where the \(WCSS\)
cannot be further reduced by moving any feature to another cluster,
despite a better solution existing
\citep[i.e., a local minimum;][]{timmerman2013}. It is, therefore often
recommended to run the k-means clustering with several different
starting positions.

In our case, we entered the participants' PC-scores from the feature
reduction step into the k-means algorithm. Because we did not know the
underlying number of clusters within our sample, we calculated the
cluster solutions for \(k=\{2, \dots , 10\}\) . To avoid local minima we
used 100 random initial centroid positions for each run. Each of the 9
cluster solutions converged within the iteration limit. In the next
step, we will then evaluate which of the extracted cluster solutions
offers the best fit with the data.

\subsection{Cluster Evaluation}
Now that the participants have been assigned to their respective clusters based on the similarity of their time series features, the final evaluation step includes two main elements, (1) evaluating the performance of the clustering analyses to choose an optimal solution and (2) interpreting the extracted clusters conceptually.

\subsubsection{Performance}
Performance evaluation often means assessing the accuracy, stability, and separation or purity of the clustering \citep{keogh2003}. Importantly, any evaluation of the results depends on the research questions, the data, and the methods used. However, broadly speaking, evaluation methods can be categorized based on whether the true cluster labels are known or not \citep{saxena2017}. If true class labels are known, the cluster assignments can be compared to the true class labels --- using measures such as the F-measure, adjusted Rand index, mutual information, and normalized mutual information \citep[i.e., external evaluation; e.g.,][]{liao2005}. However, if the true cluster assignments are unknown, as with our psychological time series, the quality of the clusters is assessed based on the characteristics of the data itself, such as separation and homogeneity of the clusters, or goodness of fit indices \citep[i.e., internal evaluation; e.g.,][]{Aghabozorgi2015}.

In our own illustration example, we used the \texttt{cluster.stats()}
function from the \texttt{fpc} \textsf{R} package, which calculates a
wide variety of internal cluster validity statistics for each of the
extracted clustering solutions. With real-world data no single
evaluation measure is likely perfect, and different measures may produce
different results depending on the characteristics of the data and the
research question being addressed \citep{kittler1998}. It is therefore
important to consider a variety of evaluation measures and to carefully
interpret the results in the context of the specific analysis
\citep{vinh2009}. We found that across most indices, the analysis with
\(k=2\) clusters performed the best. Three commonly reported indices we
would like to highlight are the comparison of within clusters sum of
squares, the average silhouette score, and the Calinski-Harabasz index.
The first statistic we looked at was the total within-cluster sum of
square \(WCSS\) (see also \equatref{eq:kWCSS}). While the within-cluster
variation will naturally decrease with (more) smaller clusters, we
observed that the decrease in \(WCSS\) was largest until \(k=2\) after
which the decrease was much smaller. This method is also sometimes
referred to as the `elbow method' \citep{syakur2018}. We then looked at
a second, commonly used measure, the average silhouette score. This
statistic measures the degree to which each time feature data point is
similar to other points within the same cluster, compared to points in
other clusters \citep{rousseeuw1987}. In our case, the \(k=2\) solution
maximized the silhouette coefficient (\(s_2=\) 0.09). Finally, the
Calinski-Harabasz index assesses the compactness and separation of the
clusters by assessing the ratio of the sum of between-clusters
dispersion and of inter-cluster dispersion for all clusters --- thus,
the higher the score the better the performances \citep{calinski1974}.
In our case, the \(k=2\) solution also showed the highest
Calinski-Harabasz index (\(CH_2=\) 16.38; a full table of all extracted
validity statistics is available in Supplemental Material
A)\footnote{It is important to note that another commonly assessed aspect of the evaluation is determining the stability and robustness of the clusters \citep{berkhin2006}. This can be assessed by evaluating the sensitivity of the clusters to different feature sets or clustering algorithms, or by using techniques such as bootstrapping to assess the uncertainty of the clusters \citep{vinh2009}. Especially when comparing different clustering algorithms one common index is the Bayesian information criterion (BIC), where a lower BIC indicates that a model is more representative of the data \citep{vandeschoot2017}.}.
In the final \(k=2\) solution the k-means analysis also assigned a
relatively even number of participants to cluster 1 (\(n_{C_1}=\) 76)
and cluster 2 (\(n_{C_1}=\) 80).

\subsubsection{Interpretation}
The interpretation of feature-based time series clustering in psychology involves understanding the meaning and implications of the obtained clusters. In order to make sense of the clustering results, we here focus on three general aspects of the results \citep{kaufman1990}. (1) Assessing differences between the clusters in the original time series features, (2) comparing the clusters based on prototype developments, (3) comparing the clusters based on between-person differences that were not included in the initial clustering.

In short, we find that the feature-based clustering discerned two
meaningfully different groups of participants. We find an adaptive group
(cluster 1) that reports higher well-being and more positive outgroup
attitudes (\textit{median}) that are also stable over time
(\textit{MAD}, \textit{MAC}) and tend to increase over the 30 day test
period (\textit{linear trend}). This group also reported consistently
more meaningful, need-fulfilling, and cooperative outgroup interactions
(\textit{median}). This group with overwhelmingly positive experiences
stands in contrast with a more detrimental group (cluster 2). This
cluster, on average, reported much less positive, less meaningful, and
less fulfilling interactions and interaction patterns (\textit{median}).
This group also reported less positive outgroup attitudes, lower
well-being, and more discrimination experiences (\textit{median}). At
the same time, for members of this detrimental cluster (cluster 2)
conditions seemed to deteriorate over time (\textit{linear trend}), and
there was generally less consistency in the experiences they were able
to have (\textit{MAC}, \textit{MAD}, \textit{edf}).

To identify these patterns, we first inspect the clusters based on the
average values of meaningful features (see
\fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We see that for
some variables the features are generally stronger in separating the
clusters. We, for example, see that the item on
`\textit{how cooperative the interaction was}' distinguishes the two
clusters across almost all seven features (except for the
\textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}). Compare
this to the `\textit{outgroup attitudes}' item where the differences
between the clusters are much smaller for almost all features. We then
inspect the clusters with a focus on the features (see
\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
variable focus, we can see more clearly that some features are better at
distinguishing the clusters across variables. For example, \textit{MAD}
and \textit{median} distinguish the two clusters across almost all
variables (except for the item of whether the interaction was
representative of the outgroup). These two features stand in stark
contrast to other features, such as the
\textit{lag-1 auto correlations}, which showed much smaller differences
between the two clusters (see \fgrref[B]{fig:clusterFeatVar}). This
offers some information on which features were most important in
understanding the two extracted groups. Taking these two perspectives
together, we can also focus on individual features or variables in
particular. We, for example, see a strong difference in the average
well-being, where participants in cluster 2 showed a much lower median
well-being over the time series. At the same time, in terms of
stability, both groups have virtually identical average \textit{MAC}
statistics for well-being (see \fgrref[A]{fig:clusterFeatVar}). There
are, thus, variables and features that distinguish the clusters better
than others and a combination of variables and features lets us explore
meaningful group differences in more detail. In our case, we see that
the central tendency, variability, and linear trend are best at
distinguishing a group with mainly positive experiences (cluster 1) from
a group with a more negative experience (cluster 2). We also see that
our clusters line up with the past literature on the importance of
focusing on simpler and more meaningful statistics
\citep{bringmann2018c, eronen2021a}.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons based on Features and Variables}
  \label{fig:clusterFeatVar}
  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb.pdf}
  \caption*{Note: \\
  "Int." = outgroup interaction. Within the "(B) Feature Focus" subplot, the 'n' and 'Discrimination' comparison variables were not part of the original time series clustering.}
\end{figure}

In the second step, we look at the prototypical trajectories of the
clusters. For k-means clustering it is often recommended to use the
average over time of the responses within the cluster
\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}. Additionally, finding cluster prototypes is often substantially easier with embedded clustering methods because in many cases a cluster-level model is estimated as part of the expectation–maximization procedure \citep[e.g.,][]{denteuling2021} or \textit{S-GIMME} \citep[e.g.][]{lane2019}. For medoid-based clustering algorithms, a common approach is simply using cluster medoid as the prototype \citep{kaufman1990}.}.
Immediately striking are the mean differences, where participants in
cluster 1 had more meaningful and fulfilling outgroup interactions and
also consistently reported more voluntary and cooperative interactions
but fewer accidental and involuntary interactions. The same cluster
(cluster 1) also reported an increase in need-fulfilling interactions
over the 30-day period and an increase in interactions that were
representative of the outgroup. Whereas the other cluster (cluster 2)
showed a decrease in voluntary, cooperative, and positive interactions
over the 30 days. This `deterioration' cluster (cluster 2) also saw a
decrease in general need fulfillment and experienced well-being over the
30 days (see \fgrref[B]{fig:clusterTs}). We also see that while
interaction representativeness, outgroup attitudes, and well-being are
relatively stable for both clusters, the deteriorating cluster (cluster
2) also showed substantially higher variability and instability on most
of the other variables (also see \fgrref[A]{fig:clusterTs}).

Finally, we can also assess the clusters across other individual
difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
comparison allows us to check for data artifacts, as well as check
whether the developmental clusters are associated with important social
markers and individual differences. To illustrate artifact checks, we
added the number of ESM measurements into the comparison and find that
the participants in the deterioration cluster (cluster 2) on average
completed slightly more ESM surveys in general and reported on more
intergroup interactions in particular (see \(n\) in
\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
ensured that the general time frame and completion rates are similar for
all participants and indeed the numbers in ESM measurements generally
are largely similar (e.g., see \(n\) for well-being and outgroup
attitudes). However, the difference in the reported number of
interactions might indicate either a clustering artifact or a meaningful
difference. The higher average number of interactions in cluster 2
could, for example, indicate a clustering artifact if variances are
substantially larger due to the larger samples
\citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
In our case, this seems less likely because one out of four variables
did not differ in terms of the MAD (i.e., our selected measurement of
the time series variance; see \fgrref{fig:cluster_comparison_mad} for an
illustration). At the same time, however, the difference in the number
of experienced interactions might also indicate a meaningful difference,
where the deteriorating cluster (cluster 2) on average reported more
outgroup interactions (\textit{difference} = 1.03, \(t\)(150.83) = 7.50,
\(p\) \textless{} .001, \textit{95\%CI} {[}0.76, 1.30{]}), but these
interactions were less voluntary (\textit{difference} = -1.04,
\(t\)(108.89) = -7.71, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.31,
-0.77{]}), less meaningful (\textit{difference} = -1.00, \(t\)(136.40) =
-7.16, \(p\) \textless{} .001, \textit{95\%CI} {[}-1.28, -0.73{]}), and
less positive (\textit{difference} = -1.38, \(t\)(152.31) = -11.94,
\(p\) \textless{} .001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while
more research is needed for a conclusive test, our data seems to suggest
that the differences in reported interactions are a meaningful
difference between the clusters. Such a finding would also be in line
with past research highlighting the role of negative intergroup
interactions in explaining intergroup relations
\citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}.

\begin{figure}[!ht] %hbtp
  \caption{Cluster comparison of Median Absolute Deviation for all variables}
  \label{fig:cluster_comparison_mad}
  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_mad.pdf}
\end{figure}

To further illustrate the utility of assessing out-of-feature individual
differences, we also compare the two samples in terms of the
participants' self-reported discrimination experiences in the
Netherlands (measured during the post-measurement). When looking at the
group comparison, we find that participants in the deteriorating cluster
(cluster 2) reported substantially higher levels of everyday
discrimination (\textit{difference} = 0.40, \(t\)(151.71) = 2.56, \(p\)
= 0.011, \textit{95\%CI} {[}0.09, 0.71{]};
\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
(e.g., the sum of specific ESM measurements) and cross-sectional
variables (e.g., general discrimination differences) that were not
included in the original clustering step can be used to explore and
understand the cluster differences in more detail.

This cluster separation, then, has a number of empirical and practical
applications. Firstly, the clusters are descriptive. With tens of
variables, hundreds of participants, and thousands of measurements,
singular descriptive statistics are often not able to capture the
complex patterns that describe the data set. The feature-based
clustering offers some direct insight into the complexity within the
data set. In our empirical example, we, for example, see that
participants are meaningfully distinguished by a combination of high
(vs.~low) central tendency, variability, and linear trend. Secondly, the
clusters identify important groups. The adaptive and deteriorating
groups offer starting points for empirical exploration as well as
practical interventions. Researchers can start probing what exactly
distinguishes the two groups further and generate new bottom-up
hypotheses. Practitioners in the resettlement field can use the group
separation to identify individuals in need of assistance and can explore
contextual factors that might contribute to the difficulties some might
face. In our illustration we, for example, found that participants in
the deteriorating cluster (cluster 2) reported less need fulfilling
interactions over time. Thirdly, the feature-based approach is flexible
and meaningful. We were able to use a wide range of time series features
that have been central in the ESM literature and were able to use them
directly to identify meaningful groups. For our empirical illustration
we, among others for example, chose to focus on whether participants
differed in their average well-being (i.e., \textit{median}), how much
their well-being would vary over time (i.e, \textit{MAD}), and whether
their well-being would on average increase or decrease over time (i.e.,
\textit{linear trend}). Alternatively, for others cyclical patterns
might be more important --- for example, whether well-being was higher
on weekends. Importantly, in any case, we did not need to translate
these dynamic features into probabilistic inference models (e.g., VAR
models) to cluster the participants.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons over time}
  \label{fig:clusterTs}
  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
  \caption*{Note: \\
  Subplot (A) displays the variable cluster means at every measurement occasion. Subplot (B) shows the GAM spline for each cluster across the measurement occasions. The thinner lines present all individual time series.}
\end{figure}

\newpage
\section{Discussion}
% aims re-iterated
The purpose of this article was to introduce feature-based time series clustering as an amenable and transparent approach to understanding between-person differences in developmental patterns of psychological time series data. Rather than relying on person-specific model parameters, which can be restrictive and assumption-bound, we argue for the more flexible and theoretically grounded approach of directly clustering on relevant features of the time series data. By leveraging the rich array of dynamic measures, our approach offers the advantages of flexibility, fewer strict assumptions, and improved interpretability, thus potentially enriching our understanding of heterogeneous psychological processes in intensive longitudinal studies.

% summary
To illustrate the practical utility of the approach, we applied the method to real-world empirical data that highlight common ESM issues of multivariate conceptualizations, structural missingness, and nonlinear trends \citep[e.g.,][]{ariens2020}. With the real-world data, we followed a stepwise approach to discuss key issues during input selection, feature extraction, feature reduction, feature clustering, and cluster evaluation. Within this step-wise approach, our article shows that feature-based clustering offers an excellent fit for psychological research, as both the time series features and the analysis steps are well established within the field, and statistical packages are readily available. Time series features (such as means or linear trends) are not only easy to extract, but also hold conceptual meaning for psychological data and can be chosen to address specific research questions (also see \tblref{tab:esmFeatures}).

Importantly, we show that feature-based clustering is not only approachable but provides interpretable and transparent insights about the grouped patterns. For our example of migration experiences, the method was useful to discern adaptive from more stressful experiences and helped to contextualize divergent experiences. We found that some variables, such as interaction quality perceptions or need fulfillment were particularly important in distinguishing the groups (see \fgrref[A]{fig:clusterFeatVar}). Similarly, we found that the central tendency (\textit{median}), variability (\textit{MAD}), and linear trend (\textit{slope}) were the most impactful dynamic features in discerning the trajectory clusters (see \fgrref[B]{fig:clusterFeatVar}). Jointly these two approaches allowed us to identify a cluster that had generally positive and improving experiences while the other cluster had more negative and deteriorating experiences. We were even able to further contextualize the results with out-of-feature comparisons, where we found that the group with the more difficult experiences also reported substantially more discrimination experiences during the post-test (see, e.g., \fgrref[B]{fig:clusterFeatVar}). In short, the feature-based approach allowed us to identify directly interpretable and meaningful groups, where we transparently know what data input the clusters are based on.

\subsection{Limitations}
While feature-based time series clustering offers a promising approach to understanding psychological time series data, it is not without limitations. In particular, feature-based clustering has both usability- and robustness limitations across its multiple steps.

In terms of convenience, each of the steps requires users to make an informed decision about the choice of method and algorithm. These additional steps of decision-making and transparency increase the initial barrier to entry. We hope that our empirical illustration, the sample code, and the custom functions, offer a relatively generalizable procedure that showcases the ease of use, but clustering, unfortunately, does not offer a universal one-size-fits-all solution.

In terms of methodological robustness, the variety of methods in each of the steps also brings with it the potential inconsistent results between methods \citep[e.g.,][]{bastiaansen2019}. A different set of variables, time series features, or a different clustering algorithm, might have resulted in substantially different cluster assignments. While the variety and diversity of methods are helpful in finding options even for more complex types of data, different algorithms often offer different results \citep[e.g.,][]{keogh2005}. And even when patterns produce robust clustering solutions across algorithms, individual methods might still have their idiosyncratic shortcomings \citep{xu2015}.

As an illustration, the choice of time series features to extract from the time series data is a critical step that can significantly influence the results of the clustering process. In the current example, we chose to extract time series features such as means, autocorrelations, and linear trends, which are psychologically and conceptually meaningful in interpreting our time series clusters. However, this selection is not exhaustive and may not capture all relevant aspects of the time series data. For example, we did not consider attributes like periodicity or spectral density, which could shed light on the data's cyclical patterns. The choice of time series features largely hinges on the researcher's specific research question and assumptions about the data, thereby injecting a level of subjectivity into the process. Similar challenges arise with the choice of the clustering algorithm or the cluster illustration. These challenges are not unique to feature-based clustering, rather they are common to all clustering approaches \citep{liao2005, horne2020}. However, it is important to remember that multi-stepped data-driven approaches are particularly vulnerable to the impact of the researchers' degrees of freedom.

One potential remedy to many of the limitations of feature-based clustering lies in transparently and reproducibly reporting the decisions of the user for each of the analysis steps. In our own description of the method, we have provided a range of options and have motivated our own choices to facilitate the transparency of the individual steps and decision moments. Beyond the structures proposed here, \citet{vandeschoot2017} have proposed an extensive checklist for latent trajectory studies. Most of their recommendations and reporting guidelines also apply to feature-based clustering and might even offer a template for researchers who hope to preregister their analysis procedures \citep[also see][]{kirtley2021}.

\subsection{Implications}
Notwithstanding the limitations, we believe that feature-based clustering offers exciting new potential for researchers and practitioners assessing psychological time series.

For researchers, the feature-based time series clustering approach offers a number of compelling implications. The flexibility and interpretability mean that feature-based time series clustering can be applied to a wide range of data types and research questions. 
The method can be used to contextualize preexisting groups by extracting their time series features and comparing a data-driven approach with existing group labels. Furthermore, the feature-based approach can also be used as an exploratory, descriptive, or predictive approach to intensive longitudinal data. By reducing the complexities of ESM data to important and meaningful patterns, a bottom-up approach can aid in the creation of more embedded theories and interventions, or simply in describing the often complex and heterogeneous data researchers collect during ESM studies.

Looking ahead, the feature-based time series clustering approach opens up new avenues for future research. While the approach has shown promise in dealing with the challenges of dimensionality, missingness, and time scales, there is potential for further refinement and expansion. To showcase the exciting potential for future methodological integrations, we will briefly consider the broad range of alternative approaches to time series clustering (see \fgrref{fig:tsClustTax}).

For instance, given that the approach does not assume the stationarity restrictions of many model-based approaches, future research can now more easily integrate many of the (non-)linear trend features. Research on capturing nonlinear trends has been growing over the past years and there are exciting possibilities to bring these developments to ESM data \citep{bringmann2023}. For example, bicoherence metrics, polynomial-, and differential equation parameters may be used to capture the type and structure of nonlinear developments \citep[e.g. shape-based approaches in \fgrref{fig:tsClustTax}; see also][]{caro-martin2018, mayor2022}. New time series features capturing nonlinear structures would add to the under-studied (non-)linear trend features of ESM data.

% Additionally, feature-based clustering algorithms traditionally do not provide a cluster-level prototype trajectory. Instead, users have to extract, often suboptimal, approximations of the group-level patterns (e.g., see the section on \textit{Cluster Evaluation}). Iterative or embedded clustering approaches, such as mixture models or group-based trajectory models, often estimate a group-based model and variance around that group-based model \citep[see \fgrref{fig:tsClustTax} and \tblref{tab:clusterApproaches}; e.g.,][]{denteuling2021, lane2019}. To provide similar functionality for feature-based models, researchers may evaluate the utility of using recurrent neural networks, such as long short-term memory (LSTM) models to generate cluster-level trajectories. One could, for example, train an LSTM based on the feature matrix and the original time series as the target \citep[e.g., see][]{fraley2002, nagin1999}. Alternatively, rule-based classifiers could extract rule sets that narrow down the conditions under which a given feature set is assigned to a specific cluster \citep[][]{benard2019}. This could potentially allow us to generate cluster-level trajectories, giving us an estimate of the typical trend for each group.

\begin{figure}[!hbtp] %hbtp
  \caption{Time Series Clustering Taxonomy}
  \label{fig:tsClustTax}
  \centering\includegraphics[width=\textwidth]{figures/TS Cluster Flow/tsClustTax.pdf}
  \caption*{Note: \\
  The taxonomy only exemplifies some of the basic differences between a number of common time series clustering approaches. As such, the taxonomy and the notes are neither exhaustive nor complete in distinguishing different approaches. Additionally, terms and labels are used inconsistently across different types of literature and are chosen to avoid overlapping labels.}
\end{figure}

Beyond the direct academic use, the feature-based time series clustering approach also addresses practical and applied uses. For practitioners, the approach offers a practical and grounded method for dealing with the challenges of complex and messy data from multiple patients, customers, or users. Not only does the approach directly deal with dimensionality, missingness, and time scales in the time series, but the interpretability and transparency aspects offer particular utility in applied settings, where the costs of misspecification are high. Additionally, the approach is also more readily accessible to practitioners who may not have extensive training in complex data analysis techniques. We provide practical algorithm overviews and readily available code for data preparation, analysis, and interpretation. The ability to identify and interpret meaningful patterns in time series data can have significant implications for practice, particularly in fields such as clinical, organizational, or social psychology, where understanding individual differences and developmental patterns can inform interventions and decision-processes.

% conclusion
In conclusion, we show that feature-based time series clustering can effectively reduce the complexities of psychological time series data to important and meaningful patterns. It does so with more flexibility, versatility, and less strict assumptions than many of the commonly used approaches to date. As such the feature-based time series clustering approach addresses key challenges in the field and aids researchers and practitioners in describing and exploring patterns across participants. We hope that the method adds to the methodological toolkit of ESM researchers and promotes the creation of more embedded methods, theories, and interventions.

% Tables
% Example
%\input{Tables/descrFullWide}

% Figures

\printbibliography

\appendix

\section{ESM Data Challenges and Promises}
\label{app:ChallengesAppendix}

\subsection{Promises}

Time series clustering has a number of conceptual use cases with psychological data. Prime among them is the ability to reduce the time, variable, and person complexity by extracting and organizing participant-level structures. These reduction and structuring qualities can be essential in detecting phenomena and extracting more abstract functional principles \citep[][]{eronen2021a}. These phenomena and principles can be meaningful differences that distinguish participants in different clusters, as well as important patterns, trends, and relationships that participants share within a cluster \citep[e.g.,][]{schrodt2000}. Once distinct groups and patterns have been identified, researchers can examine the extent to which these within-group and between-group structures are associated with other variables of interest, such as personality traits, demographic characteristics, or other psychological constructs \citep[e.g.,][]{monden2022}. By detecting meaningful and robust structures and patterns, time series clustering can, thus, be used to inform the development of robust theories as well as targeted interventions and therapies for individuals, for example, with mood disorders and other psychological conditions \citep[e.g.,][]{borsboom2021, eronen2020}.

However, while clustering can be incredibly useful, arriving at these clusters critically depends on two core challenges. First, time series need to be made comparable in order to identify key (dis)similarities and second, comparable (dis)similarities need to be accurately distinguishing into different groups \citep[e.g.,][]{Aghabozorgi2015}. In practice, most psychological time series cannot be compared based on the raw data itself. This is the case because in most cases the raw time series include too many data points --- sometimes referred to as the dimensionality curse \citep[e.g.,][]{altman2018} --- and, more importantly, individual time points are oftentimes not directly comparable between participants in psychological data and would lead to misspecifications \citep[e.g., ][]{faloutsos1994}. While such issues can be avoided with transformations for highly regular, controlled, and comparable time series such as EEG data \citep[e.g.,][]{huang1985}, most ESM researchers are usually not interested in directly comparing individual timepoints between participants but are interested in developmental patterns and relationships.

As a result, most psychological time series are summarized via a numerical representation and these numerical summaries are then comparable and used to cluster participants (e.g., \citealp[]{timmerman2013}; see \fgrref{fig:tsClustTax}). Ideally, the representations that summarize the original time series data should (1) capture the original data accurately without loosing too much information, and (2) should be conceptually meaningful \citep[][]{vandermaaten2009}. Extracting accurate and meaningful representations of the time series can be essential for understanding what goes into the clustering algorithm (i.e., assists with explainability) and can be crucial in making sense of the final cluster output \citep[i.e., assists with interpretability; e.g.,][]{Kennedy2021}.

\subsection{Challenges}

We will briefly consider which challenges modern ESM data introduce and what qualities are called for in an extension of the clustering repertoire. We particularly highlight issues of dimensionality, non-equidistant or missing measurements, an interest in non-stationary trends, as well as inconsistent/diverse time scales.

Concerning dimensionality issues, especially more abstract psychological experiences often need a wider variety of measurements to be captured adequately. Today, few clinical conditions are captured with a single symptom measure \citep[e.g.,][]{cramer2016}, emotions are rarely assessed in isolation \citep[e.g.,][]{reitsema2022}, and socio-cultural experiences are now widely considered to be multimodal \citep[e.g.,][]{Kreienkamp2022d}. This also means that modern analysis techniques increasingly need be able to accommodate an increased focus on multivariate developments. At the same time, however, an increase in the number of considered variables tends to come at the expense of computational load for model estimations, and clustering models may not converge \citep[the aforementioned dimensionality curse;][]{altman2018}. A modern time series clustering technique should consequently be able to summarize and structure multivariate phenomena without running into computational load issues.

Another common type of data are measurement regiments that collect data in irregular time intervals (i.e., non-equidistant measurements). Common are, for example, procedures where participants are asked to respond at random times throughout the day (i.e., signal-contingent) or following specific natural events of interest \citep[i.e., event-contingent; see][]{shiffman2008, myin-germeys2018}. Under such conditions data tends to violate the equidistance assumption that is expected by many time series models \citep[][]{hamaker2017}. Smaller issues of non-equidistant data can be avoided with transformations \citep[e.g., dynamic time warping,][]{berndt1994} or newer modeling procedures \citep[e.g., continuous-time models;][]{dehaan-rietdijk2017} but for many analyses, including some cluster approaches, non-equidistant measurements remain a prevalent issue.

Structural missingness remains an even more strenuous challenge. Structural missingness occurs when data is missing because it logically cannot be collected \citep[as opposed to probabilistically missing data;][]{little2020, mclean2017}. Often, however, researchers might want to include variables in their models that are not available under all conditions. Follow-up and event-contingent questions are a common example in ESM studies. Researchers, for example, ask about the frequency, intensity, or duration of symptoms --- but only if a symptom was present \citep{kivela2022}. Such approaches become specifically critical in cases of sensitive questions such as questions about suicidal ideation or other potentially trauma-inducing questions \citep[e.g.,][]{glenn2022}. The most common practice for structurally missing data is to either exclude the variable or any measurement that has no structurally missing data \citep[e.g.,][]{lavori2008}\footnote{This is the case because the most commonly used models require complete data \citep{schafer2002} and structurally missing data cannot be imputed as it logically does not exist \citep[e.g.,][]{lavori2008}.} --- neither option suits a research question that wishes to include variables with common structural missingness, such as event-specific or follow-up questions. In short, new clustering approaches should be able to deal with structurally missing data in order to address modern ESM data.

When it comes to studying developmental trajectories, psychological researchers are often also interested in nonstationary processes because they are more representative of the complex, dynamic patterns of the human mind. In psychology, nonstationary processes are typically used to study phenomena such as cognitive development \citep[][]{quartz1997}, decision-making \citep[][]{ratcliff2016}, and emotion dynamics \citep[][]{bringmann2018b}. These processes are often characterized by changes in the underlying statistical properties of the data over time, such as changes in the mean or variance \citep[][]{molenaar2009}. Especially when considering changes in mean levels, researchers are often interested in nonlinear changes because they describe human functioning better. For example, in decision making people might switch between choices \citep[][]{ratcliff2016}, or patients reducing medication might experience mood swings \citep[][]{helmich2020a}. Similarly, psychologists are often also interested in how variances change over time. This is especially the case because several changes in an individual's variance have been found to be indicative of critical changes, including depression relapses and symptom shifts more generally \citep[e.g.,][]{schreuder2020, wichers2020}. There is, thus, also a need for time series clustering algorithms that capture nonstationary processes, including nonlinear trends.

Psychological time series often exhibit complex patterns and relationships that can change over different time scales. For example, a time series of daily mood ratings may show a weekly pattern, with higher ratings on the weekends and lower ratings during the week. At the same time, the series may also exhibit a longer-term trend, with overall mood levels increasing or decreasing over the course of several months or years \citep[e.g.,][]{Ram2014}. These different time scales can be studied separately or in combination, using different statistical techniques and modeling approaches \citep[][]{bertenthal2007, jeronimus2019a}. Different time scales can become an even more difficult issue when different variables in a model develop on different time scales \citep{bringmann2022b}. Different time scales are thus also a concern clustering approaches should be able to address.

It is this background of the common challenges of current ESM data, upon which we propose to consider feature-based clustering. The flexibility of using a wide variety of time series features that represent the important developmental patterns allows users to circumvent many of the issues with multi-dimensionality, non-equidistant or missing measurements, non-stationary trends, as well as diverse time scales.

%These shortcomings, however, stand in sharp contrast with the types of data researchers commonly collect and do not align with common research interests in the field. Psychological researchers might, for example, collect data based on the occurrence of natural events, which tends to result in non-equidistant measurements \citep[e.g.,][]{myin-germeys2018, hamaker2017} and context-specific missingness that cannot be imputed, for instance, interaction quality perceptions \citep[e.g.,][]{kivela2022, lavori2008}. At the same time, researchers are often specifically interested in non-stationary developments, looking, for example, at how symptom levels and -variations change over time --- often abruptly and non-linearly \citep[e.g.,][]{bringmann2018b, helmich2020a}.

\end{document}