### 2024-11-11 latex-results/03FeatureExtraction.tex

\paragraph{Central tendency.}

The central tendency refers to the statistical measures that represent
the ``typical'' or ``average'' of a set of data. The most common
measures of central tendency are the mean, median, and mode
\citep{weisberg1992}. As a familiar statistic from probability theory,
the central tendency sits at the heart of many fundamental questions
about psychological time series. Researchers might, for example, be
interested in whether ``Over a one-month period, are some people happier
than others?''

For the central tendency feature of our illustration, we chose the
median (\(M\)), which effectively addresses potential complications
arising from non-normally distributed responses or outliers within time
series datasets \citep{weisberg1992}. To compute the median, it is
imperative to differentiate between two types of time series
representations for a given variable \(j\) related to participant \(i\):
the chronological series and the ordered series. The chronological time
series, denoted by \(X_{ij}\), encapsulates the sequence of observations
\(\{x_{ij1}, x_{ij2}, ..., x_{ijn}\}\) for variable \(j\) concerning
participant \(i\), organized by their temporal occurrence. Here,
\(x_{ijt}\) signifies a specific observation at time \(t\) within this
sequence. In contrast, the ordered time series, represented as
\(\mathbf{X}_{ij}\), is derived from \(X_{ij}\) by sorting the
observations in ascending order of magnitude. This ordered set is
expressed as
\(\{\mathbf{x}_{ij1}, \mathbf{x}_{ij2}, ..., \mathbf{x}_{ijn}\}\), with
each \(\mathbf{x}_{ijk}\) corresponding to the \(k\)-th element in the
reordered series \(\mathbf{X}_{ij}\).

The median \(M(\mathbf{X}_{ij})\) is then the value located precisely at
the center of the ordered time series \(\mathbf{X}_{ij}\). Depending on
whether the total number of observations (\(n\)) is odd or even, the
median is either the middel \(k\)-th element if \(n\) is odd, or the
average of the two middle values if \(n\) is even:

\begin{equation} \label{eq:median}
  M(\mathbf{X}_{ij}) = 
  \begin{cases}
    \mathbf{x}_{ij\left(\frac{n+1}{2}\right)} & \text{if $n$ is odd} \\
    \frac{\mathbf{x}_{ij\left(\frac{n}{2}\right)} + \mathbf{x}_{ij\left(\frac{n}{2} + 1\right)}}{2} & \text{if $n$ is even}
  \end{cases}
\end{equation}

This approach ensures that the median is a reliable indicator of central
tendency in time series analysis, unaffected by data distribution
asymmetries or the presence of outliers.

\paragraph{Variability.}

Variability captures the degree to which a set of data differs from the
central tendency and is sometimes also referred to as the dispersion or
spread of the data \citep{weisberg1992}. Common measurements of the
variability are the variance or standard deviation as well as their
robust counter parts. In time series analyses, variability is
conceptually important because information about the distribution and
diversity of the data has been found to be indicative of worse
psychological states \citep{myin-germeys2018, helmich2021}. Person-level
differences of ESM measurements have, for example, been associated with
higher levels of psycho-pathological recurrences among depression
patients \citep{timm2017}. As such, psychological researchers and
practitioners are often empirically interested in between-person
differences in variability. Researchers on polarization and
radicalization might for example ask: ``Are people settled in their
attitudes towards migrants or do they vary across the measurement
period?''

For our illustration data, we chose the
\textit{Median Absolute Deviation} (\(MAD\)) to gauge the variability
within our time series data. This choice is motivated by the robustness
of MAD, particularly its resilience to the effects of non-normal
distributions and outliers, which can significantly skew traditional
variability measures like the standard deviation \citep{weisberg1992}.
For a given variable \(j\) and participant \(i\), the MAD is calculated
by first determining the median (\(M\)) of the ordered time series
\(\mathbf{X}_{ij}\) as outlined in \equatref{eq:median}. We then compute
the absolute deviations of each observation in the time series
\(X_{ij}\) from this median value. Specifically, for each time point
\(t\), we calculate the absolute difference between \(x_{ijt}\) and the
series median \(M(\mathbf{X}_{ij})\). The MAD is then the median of
these absolute deviations:

\begin{subequations}\label{eq:mad}
    \begin{align} 
      MAD(X_{ij}) &= M(\left| x_{ijt} - M(\mathbf{X}_{ij}) \right|) \label{eq:mad_general} \\
                  &= M(\{ |x_{ij1} - M(\mathbf{X}_{ij})|, |x_{ij2} - M(\mathbf{X}_{ij})|, \ldots, |x_{ijn} - M(\mathbf{X}_{ij})| \}) \label{eq:mad_detailed}
    \end{align}
\end{subequations}

The calculation of MAD focuses on the magnitudes of deviations, ensuring
it provides a robust measure of dispersion that reflects the inherent
variability in the time series data.

\paragraph{Instability.}

Instability captures the average change between two consecutive
measurements \citep{ebner-priemer2009, jahng2008}. While instability is
conceptually related to the variability feature, variability does not
take into account temporal dependency, whereas instability looks at the
`jumpy-ness' of the data over time. In other words, variability reflects
the range or diversity of values in the un-ordered time series data,
while instability reflects the fluctuation or inconsistency in a time
series data over time \citep{trull2008, houben2015, koval2013}. For
example, if a person has rapid and extreme changes in mood their mood is
highly unstable, while if a person's mood responses span a wide range
over the entire study period, their mood is highly variable
\citep[note that this does not need to be rapidly changing or instable, e.g., when there is linear increase over time; also see][]{jahng2008}.
Within psychological time series, instability measurements have
especially been important in the research of borderline personality
disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also
in understanding early warning signals more generally
\citep{wichers2019}. Conceptually, the instability feature, thus,
relates to a broad range of research questions, including: ``What is the
nature of the identification changes in those who start working in a new
country?'' or ``Do strong daily fluctuations in self-esteem reflect the
process of identity formation in adolescents?''

For our data we chose the \textit{mean absolute change}
\citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at
the average absolute difference of two consecutive measurements \(x\) at
time points \(t\) and \(t-1\), for each time series \(X\) of participant
\(i\) and variable \(j\).

\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=2, \ldots, t}\left|x_{t}-x_{t-1}\right|
\end{equation}

Another common measurement of instability is the
\textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is
often preferred where differences in magnitude are more important than
the frequency of those changes, for example, when big shifts in time
series are considered more impactful or when outliers are meaningful and
need to be taken into account \citep{chatfield2003, bos2019}. For
psychological ESM data, some research suggests that amplitude and
frequency might predict different health outcomes and can be
investigated jointly \citep{wang2012a, jahng2008}.

\paragraph{Temporal dependence.}

Temporal dependence in time series data refers to the degree to which a
time series is influenced by its past values, exhibiting patterns of
behavior that may be regular over different time scales
\citep{dmello2021}. In the context of psychological time series, an
important aspect of temporal dependence is \textit{inertia} --- how much
a measurement carries over to its next measurement
\citep{kuppens2010, suls1998}. If inertia is high a development tends to
stay in a certain state. Because high inertia is resistant to change, in
emotion dynamics high inertia of negative affect has been found to be
indicative of under-reactive systems and to be characteristic of
psychological maladjustment \citep{kuppens2010}. In a similar vein, high
inertia in negative affect at baseline was predictive of the initial
onset of depression \citep{kuppens2012}. Conceptually, inertia is more
broadly connected to research questions such as: ``Do patients stay in a
negative mood for several measurements?'' or ``Do migrants stay with
their language practice for several days at a time?''

For our illustration case, we chose the commonly used autocorrelation or
autoregression with a lag-1 to capture the inertia. High autocorrelation
values can indicate high levels of inertia, while low autocorrelation
values may indicate a more unpredictable or volatile time series
\citep{dejonckheere2019}. The lag--1 autocorrelation \(r_{ij,1}\) looks
at the average correlation between a measurement \(x\) and the preceding
measurement \(x_{t-1}\) for the time series \(X\) of participant \(i\)
and variable \(j\) with \(n\) measurements.

\begin{equation} \label{eq:ar}
  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
\end{equation}

Where \(\overline{x}_{ij}\) is the mean of the time series \(x_{ij}\),
calculated as:

\begin{equation} \label{eq:mean_for_ar1}
  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
\end{equation}

While inertia captures the simplest case of temporal dynamics, lag-1, we
acknowledge that temporal dependence in psychological time series may
also exhibit more complex relationships, including higher lagged auto
correlations or cyclical relationships (fourier coefficients, or
continuous wavelet transforms are often used to capture such
relationships).

\paragraph{Linear trend.}

In non-stationary time series, a linear trend can be observed when there
is a consistent increase or decrease in the data over time
\citep{nyblom1986}. For psychological time series, researchers have, for
example, pointed out the importance of linear trends in interpersonal
communications \citep{vasileiadou2014}, and emotion dynamics
\citep{oravecz2016}. Theoretically, linear trends are often considered
the simplest way of assessing whether a psychological theory of change
is appropriate \citep{gottman1969}. In empirical practice, linear trends
are, thus, commonly exemplified by research questions such as ``Do
patient symptoms improve consistently?'' or ``Does worker productivity
decline continuously?''

For the variables in our illustration data set, we chose an overall
linear regression slope to capture the linear trend. The regression
slope \(b_{ij}\) provides the average change from one time point \(t\)
to the next across all measurements \(x\) of a time series \(X\) of
participant \(i\) and variable \(j\). The specific form of the OLS slope
formula we provide below calculates \(b_{ij}\) as the sum across all
time points of the product of the deviation of time \(t\) from its mean
\(\overline{t}\) and the deviation of \(x_{ij}\) from its mean
\(\overline{x}_{ij}\) at each time point, divided by the sum across all
time points of the square of the deviation of time from its mean
(\(\sum(t-\overline{t})^2\)). Intuitively, the formula captures the rate
of change of variable \(x_{ij}\) with respect to time. This slope will
indicate how the variable \(x_{ij}\) changes over time, controlling for
its mean value and the mean of time. If the slope is positive,
\(x_{ij}\) increases over time; if it's negative, \(x_{ij}\) decreases
over time.

\begin{equation} \label{eq:lin}
  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
\end{equation}

\paragraph{Nonlinearity.}

Changes in psychology are not always linear, instead, nonlinearity is a
common feature of psychological time series \citep{hayes2007}. As an
example, episodic disorders, such as depression, are often best
described as non-linear systems \citep{hosenfeld2015}. Similarly,
patients in recovery from depression showed sudden changes in the
improvement of depression \citep{helmich2020a}. But also substance abuse
\citep{boker1998} or attitude changes rarely develop linearly
\citep{vandermaas2003}. Conceptually, researchers might have research
questions about the type of the development: ``Is the development of
well-being a nonlinear process?'' as well as the shape and structure of
the development: ``How many spikes in well-being did a migrant
experience?''

We summarized the nonlinear trend with the
\textit{estimated degrees of freedom} of an empty GAM spline model. The
\(edf\) summarizes the \textit{wiggliness} of a spline trend line
\citep{wood2017, bringmann2017}. The degrees of freedom of a spline
model are primarily determined by the number of knots and the order of
the spline. For instance, a cubic spline with \(k\) knots has \(k\)+3
degrees of freedom
\citep{faraway2016, haslbeck2021a, castro-alvarez2024}. However, in a
penalized spline framework, which is commonly used for GAMs, the
effective degrees of freedom can be less than \(k\)+3. This is because
the model employs a smoothing parameter to control the trade-off between
the complexity (flexibility) of the model and its fit to the data,
thereby penalizing overly complex models and potentially reducing the
effective degrees of freedom \citep{marx1998}. Intuitively then an edf
of 1 would be equivalent to a linear relationship (i.e., one linear
slope parameter), whereas a higher edf (particularly an edf
\textgreater{} 2) is indicative of a non-linear trend. The estimated
degrees of freedom are commonly based on a concept called `effective
degrees of freedom' and can be represented as the trace \(tr\) (i.e.,
the sum of the diagonal elements) of the smoother matrix \(S\), a
symmetric matrix that maps from the raw data to the smooth estimates
\citep{wood2017}.

\begin{equation} \label{eq:df}
  edf = tr(S)
\end{equation}

\paragraph{Additional considerations.}

Beyond our main features of interest, we also extracted the
participant's number of completed ESM measurements to ensure that the
clusters are comparable in that regard (i.e., to exclude spurious
explanations for the cluster assignments). After the feature extraction,
we found that about 1.40\% of the extracted features are missing across
the 72 features per participant. This might, for example, happen if
participants do not have two subsequent measurements with outgroup
interactions, so that an autocorrelation with lag-1 cannot be calculated
for the contact-specific variables. The small number of missing values
indicates that the feature-based approach indeed largely avoids the
structural missingness issue. However, even the few missing values can
be an issue for some feature reduction or feature clustering algorithms.
We, thus, impute missing feature values via predictive mean matching
(PMM) with the MICE package in R, employing a single imputation and
specifying a maximum of 50 iterations and a fixed seed for convergence
and reproducibility \citep[][]{buuren2011}. We chose PMM for its ability
to preserve the original data distribution without assuming normality
and robustly handling multiple data types \citep{vanbuuren2006}. Note
again that with this procedure we only need to impute an extremely small
number of missing values as most feature calculations can use the
available data instead.