### 2024-11-11 latex-results/07ClusterEvaluationInterpretation.tex

In short, we find that the feature-based clustering discerned two
meaningfully different groups of participants. We find an adaptive group
(cluster 1) that reports higher well-being (\textit{median}:
\textit{difference} = -0.52, \(t\)(153.87) = -3.34, \(p\) = 0.001,
\textit{95\%CI} {[}-0.82, -0.21{]}; also see
\fgrref[A]{fig:cluster_comparison_features}) and more positive outgroup
interactions (\textit{median}: \textit{difference} = -1.38,
\(t\)(152.31) = -11.94, \(p\) \textless{} .001, \textit{95\%CI}
{[}-1.61, -1.15{]}), which are also stable over time (\textit{MAC}:
\textit{difference} = 0.54, \(t\)(153.98) = 3.49, \(p\) \textless{}
.001, \textit{95\%CI} {[}0.23, 0.84{]}) and tend to increase more over
the 30 day test period (\textit{linear trend}: \textit{difference} =
-0.55, \(t\)(149.90) = -3.55, \(p\) \textless{} .001, \textit{95\%CI}
{[}-0.85, -0.24{]}; also see
\fgrref[C]{fig:cluster_comparison_features}). This group also reported
consistently more meaningful (\textit{median}: \textit{difference} =
-1.00, \(t\)(136.40) = -7.16, \(p\) \textless{} .001, \textit{95\%CI}
{[}-1.28, -0.73{]}), need-fulfilling (\textit{median}:
\textit{difference} = -0.99, \(t\)(135.30) = -7.17, \(p\) \textless{}
.001, \textit{95\%CI} {[}-1.26, -0.72{]}), and cooperative outgroup
interactions (\textit{median}: \textit{difference} = -1.33,
\(t\)(120.36) = -11.28, \(p\) \textless{} .001, \textit{95\%CI}
{[}-1.56, -1.10{]}). This group with overwhelmingly positive experiences
stands in contrast with a more detrimental group (cluster 2). This
cluster, on average, reported much less positive, less meaningful, and
less fulfilling interactions and interaction patterns (\textit{median}).
This group also reported less positive outgroup attitudes, lower
well-being, and more discrimination experiences (\textit{median}). At
the same time, for members of this detrimental cluster (cluster 2)
conditions seemed to deteriorate over time (\textit{linear trend}), and
there was generally less consistency in the experiences they were able
to have (\textit{MAC}, \textit{MAD}, \textit{edf}; also see
\fgrref[]{fig:cluster_comparison_features}; for a full and interactive
comparison of all features see \situtorial).

To identify these patterns, we first inspect the clusters based on the
average values of meaningful features (see
\fgrref[A]{fig:clusterFeatVar}; \citealp{Kennedy2021}). We see that for
some variables the features are generally stronger in separating the
clusters. We, for example, see that the item on
`\textit{how cooperative the interaction was}' distinguishes the two
clusters across almost all seven features (except for the
\textit{auto-correlation}, see \fgrref[A]{fig:clusterFeatVar}). Compare
this to the `\textit{outgroup attitudes}' item where the differences
between the clusters are much smaller for almost all features. We then
inspect the clusters with a focus on the features (see
\fgrref[B]{fig:clusterFeatVar}). While this is the same data as for the
variable focus, we can see more clearly that some features are better at
distinguishing the clusters across variables. For example, \textit{MAD}
and \textit{median} distinguish the two clusters across almost all
variables (except for the item of whether the interaction was
representative of the outgroup). These two features stand in stark
contrast to other features, such as the \textit{lag-1 auto correlations}
or the \textit{GAM edf}, which showed much smaller differences between
the two clusters (see \fgrref[B]{fig:clusterFeatVar} and
\fgrref[]{fig:cluster_comparison_features}; please note that we offer
readers an interactive tool to assess the cluster differences for all
features in \situtorial). This offers some information on which features
were most important in differentiating the two extracted groups but also
shows that with real-world data, not all features will have enough range
to distinguish people on all variables (e.g., see the non-linearity
patterns in \fgrref[]{fig:cluster_comparison_features}; for a more
direct illustration of GAM edf differences see \citealp{bringmann2017}).
Taking these two perspectives together, we can also focus on individual
features or variables in particular. We, for example, see a strong
difference in the average well-being, where participants in cluster 2
showed a much lower median well-being over the time series
(\textit{difference} = -0.52, \(t\)(153.87) = -3.34, \(p\) = 0.001,
\textit{95\%CI} {[}-0.82, -0.21{]}). At the same time, in terms of
well-being stability, both groups have virtually identical average
\textit{MAC} statistics for well-being (\textit{difference} = -0.01,
\(t\)(153.96) = -0.04, \(p\) = 0.968, \textit{95\%CI} {[}-0.32, 0.31{]};
also see \fgrref[A]{fig:clusterFeatVar}). There are, thus, variables and
features that distinguish the clusters better than others and a
combination of variables and features lets us explore meaningful group
differences in more detail. In our case, we see that the central
tendency, variability, and linear trend are best at distinguishing a
group with mainly positive experiences (cluster 1) from a group with a
more negative experience (cluster 2). We also see that our clusters line
up with the past literature on the importance of focusing on simpler and
more meaningful statistics
\citep{bringmann2018c, eronen2021a, dejonckheere2019}.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons based on Features and Variables}
  \label{fig:clusterFeatVar}
  \centering\includegraphics[width=\textwidth]{figures/clusterFeatVarComb_tutorial.pdf}
  \caption*{Note: \\
  "Int." = outgroup interaction, "mad" = median absolute deviation, "mac" = mean absolute change, "lin" = linear slope, "edf" = estimated degrees of freedom of an empty GAM spline model, "ar01" = lag-1 autocorrelation, "OCC"/"occ" = out-of-cluster comparison\\
  Within the "(B) Feature Focus" subplot, the 'n (within ooc)' is an out-of-cluster comparison of the within-person available measurements for each variable; the 'between ooc (mean)' are also out-of-cluster comparisons but on a between person level. 'Measurements removed' is the person-specific count of measurement occasions removed during the missingness handling and 'Discrimination' is the scale mean of daily discimination experiences (measured during the final survey).}
\end{figure}

In the second step, we look at the prototypical trajectories of the
clusters. For k-means clustering it is often recommended to use the
average over time of the responses within the cluster
\citep[see \fgrref{fig:clusterTs};][]{niennattrakul2007}\footnote{It is important to note, however, that direct comparability can be a concern, and often times some subset selection or nonlinear alignment is necessary \citep[e.g.,][]{gupta1996}.}.
Immediately striking are the mean differences, where participants in
cluster 1 had more meaningful and fulfilling outgroup interactions and
also consistently reported more voluntary and cooperative interactions
but fewer accidental and involuntary interactions. The same cluster
(cluster 1) also reported an increase in need-fulfilling interactions
over the 30-day period and an increase in interactions that were
representative of the outgroup. Whereas the other cluster (cluster 2)
showed a decrease in voluntary, cooperative, and positive interactions
over the 30 days. This `deterioration' cluster (cluster 2) also saw a
decrease in general need fulfillment but not experienced well-being over
the 30 days (see \fgrref[C]{fig:cluster_comparison_features}). We also
see that while interaction representativeness, outgroup attitudes, and
well-being are relatively stable for both clusters, the deteriorating
cluster (cluster 2) also showed substantially higher variability and
instability on most of the other variables (although these effects are
much smaller; see \fgrref[A]{fig:clusterTs}).

Finally, we can also assess the clusters across other individual
difference variables \citep[e.g.,][]{monden2022}. This out-of-feature
comparison allows us to check for data artifacts, as well as check
whether the developmental clusters are associated with important social
markers and individual differences. To illustrate artifact checks, we
added the number of ESM measurements into the comparison and find that
the participants in the deterioration cluster (cluster 2) on average
completed slightly more ESM surveys in general and reported on more
intergroup interactions in particular (see \(n\) in
\fgrref[B]{fig:clusterFeatVar}). In our data exclusion procedures, we
ensured that the general time frame and completion rates are similar for
all participants and indeed the numbers in ESM measurements generally
are largely similar (e.g., see \(n\) for well-being and outgroup
attitudes). However, the difference in the reported number of
interactions might indicate either a clustering artifact or a meaningful
difference. The higher average number of interactions in cluster 2
could, for example, indicate a clustering artifact if variances are
substantially larger due to the larger samples
\citep[e.g., restriction of range in the smaller sample][]{kogan2006}.
In our case, this seems less likely because one out of four variables
did not differ in terms of the MAD (i.e., our selected measurement of
the time series variance; see \fgrref{fig:cluster_comparison_features}
for an illustration). At the same time, however, the difference in the
number of experienced interactions might also indicate a meaningful
difference, where the deteriorating cluster (cluster 2) on average
reported more outgroup interactions (\textit{difference} = 1.03,
\(t\)(150.83) = 7.50, \(p\) \textless{} .001, \textit{95\%CI} {[}0.76,
1.30{]}), but these interactions were less voluntary
(\textit{difference} = -1.04, \(t\)(108.89) = -7.71, \(p\) \textless{}
.001, \textit{95\%CI} {[}-1.31, -0.77{]}), less meaningful
(\textit{difference} = -1.00, \(t\)(136.40) = -7.16, \(p\) \textless{}
.001, \textit{95\%CI} {[}-1.28, -0.73{]}), and less positive
(\textit{difference} = -1.38, \(t\)(152.31) = -11.94, \(p\) \textless{}
.001, \textit{95\%CI} {[}-1.61, -1.15{]}). Thus, while more research is
needed for a conclusive test, our data seems to suggest that the
differences in reported interactions are a meaningful difference between
the clusters. Such a finding would also be in line with past research
highlighting the role of negative intergroup interactions in explaining
intergroup relations \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. A
related validity check was the inclusion of the missingness handling,
where we compared the two clusters on the average number of measurements
removed as part of the missingness handling. We find that the clusters
did not significantly differ on this metric suggesting that the
missingness handling did not affect the cluster separation (also see
\appref[]{app:ValidationAnalyses} and \situtorial).

\begin{figure}[!ht] %hbtp
  \caption{Comparison Cluster Differences by Features and Variables.}
  \label{fig:cluster_comparison_features}
  \centering\includegraphics[width=\textwidth]{figures/feature_comparison_combined.pdf}
  \caption*{Note:\\
  The figure shows the differences between the clusters in the standardized features that were entered into the dimensionality reduction (for each input variable). We display the median (panel A), the median absolute deviation (MAD, panel B), the univariate linear slope (panel C), as well as the estimated degrees of freedom of the generalized additive model splines (GAM edf, panel D). Please also note that as part of \situtorial, we provide readers with an interactive selection tool to compare cluster differences on all variables and features.}
\end{figure}

To further illustrate the utility of assessing out-of-feature individual
differences, we also compare the two samples in terms of the
participants' self-reported discrimination experiences in the
Netherlands (measured during the post-measurement). When looking at the
group comparison, we find that participants in the deteriorating cluster
(cluster 2) reported substantially higher levels of everyday
discrimination (\textit{difference} = 0.40, \(t\)(151.71) = 2.56, \(p\)
= 0.011, \textit{95\%CI} {[}0.09, 0.71{]};
\fgrref[B]{fig:clusterFeatVar}). Thus, both intensive longitudinal
(e.g., the sum of specific ESM measurements) and cross-sectional
variables (e.g., general discrimination differences) that were not
included in the original clustering step can be used to explore and
understand the cluster differences in more detail.

This cluster separation, then, has a number of empirical and practical
applications. Firstly, the clusters are descriptive. With tens of
variables, hundreds of participants, and thousands of measurements,
singular descriptive statistics are often not able to capture the
complex patterns that describe the data set. The feature-based
clustering offers some direct insight into the complexity within the
data set. In our empirical example, we, for example, see that
participants are meaningfully distinguished by a combination of high
(vs.~low) central tendency, variability, and linear trend. Secondly, the
clusters identify important groups. The adaptive and deteriorating
groups offer starting points for empirical exploration as well as
practical interventions. Researchers can start probing what exactly
distinguishes the two groups further and generate new bottom-up
hypotheses. Practitioners in the resettlement field can use the group
separation to identify individuals in need of assistance and can explore
contextual factors that might contribute to the difficulties some might
face. In our illustration we, for example, found that participants in
the deteriorating cluster (cluster 2) reported less need fulfilling
interactions over time. Thirdly, the feature-based approach is flexible
and meaningful. We were able to use a wide range of time series features
that have been central in the ESM literature and were able to use them
directly to identify meaningful groups. For our empirical illustration
we, among others for example, chose to focus on whether participants
differed in their average well-being (i.e., \textit{median}), how much
their well-being would vary over time (i.e, \textit{MAD}), and whether
their well-being would on average increase or decrease over time (i.e.,
\textit{linear trend}). Alternatively, for others cyclical patterns
might be more important --- for example, whether well-being was higher
on weekends. Importantly, in any case, we did not need to translate
these dynamic features into probabilistic inference models (e.g., VAR
models) to cluster the participants.

\begin{figure}[!ht] %hbtp
  \caption{Cluster Group Comparisons over time}
  \label{fig:clusterTs}
  \centering\includegraphics[width=\textwidth]{figures/clusterTsComb.pdf}
  \caption*{Note: \\
  Subplot (A) displays the variable cluster means at every measurement occasion. The thinner lines present all individual time series. Subplot (B) shows the GAM spline for each cluster across the measurement occasions. The thinner lines present all individual GAM Splines.}
\end{figure}