### 2024-11-11 sections/03FeatureExtraction.Rmd

---
output: latex_fragment
bibliography: ../referencesZotero.bib
csl: ../apa.csl
---

```{r}
#| label: section setup
#| include: false

# Global Chunk Options
knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  fig.path = "../figures/",
  include = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
#| label: featMiss

featureExclusion <- c(
  "mean",
  "sd",
  "rmssd",
  "ar02",
  "edf_ar",
  "_n"
)

pMissFull <- pMissFeat(
  all = featFull$features %>% select(-ends_with(featureExclusion)),
  contact = featFullContact$features,
  nocontact = featFullNoContact$features,
  title = "Feature-wise Missingess Across all Studies"
)
```

\paragraph{Central tendency.}
The central tendency refers to the statistical measures that represent the "typical" or "average" of a set of data. The most common measures of central tendency are the mean, median, and mode \citep{weisberg1992}. As a familiar statistic from probability theory, the central tendency sits at the heart of many fundamental questions about psychological time series. Researchers might, for example, be interested in whether "Over a one-month period, are some people happier than others?"

For the central tendency feature of our illustration, we chose the median (\(M\)), which effectively addresses potential complications arising from non-normally distributed responses or outliers within time series datasets \citep{weisberg1992}. To compute the median, it is imperative to differentiate between two types of time series representations for a given variable \( j \) related to participant \( i \): the chronological series and the ordered series. The chronological time series, denoted by \( X_{ij} \), encapsulates the sequence of observations \( \{x_{ij1}, x_{ij2}, ..., x_{ijn}\} \) for variable \( j \) concerning participant \( i \), organized by their temporal occurrence. Here, \( x_{ijt} \) signifies a specific observation at time \( t \) within this sequence. In contrast, the ordered time series, represented as \( \mathbf{X}_{ij} \), is derived from \( X_{ij} \) by sorting the observations in ascending order of magnitude. This ordered set is expressed as \( \{\mathbf{x}_{ij1}, \mathbf{x}_{ij2}, ..., \mathbf{x}_{ijn}\} \), with each \( \mathbf{x}_{ijk} \) corresponding to the \(k\)-th element in the reordered series \( \mathbf{X}_{ij} \).

The median \( M(\mathbf{X}_{ij}) \) is then the value located precisely at the center of the ordered time series \( \mathbf{X}_{ij} \). Depending on whether the total number of observations (\( n \)) is odd or even, the median is either the middel \(k\)-th element if \( n \) is odd, or the average of the two middle values if \( n \) is even:

```{=tex}
\begin{equation} \label{eq:median}
  M(\mathbf{X}_{ij}) = 
  \begin{cases}
    \mathbf{x}_{ij\left(\frac{n+1}{2}\right)} & \text{if $n$ is odd} \\
    \frac{\mathbf{x}_{ij\left(\frac{n}{2}\right)} + \mathbf{x}_{ij\left(\frac{n}{2} + 1\right)}}{2} & \text{if $n$ is even}
  \end{cases}
\end{equation}
```

This approach ensures that the median is a reliable indicator of central tendency in time series analysis, unaffected by data distribution asymmetries or the presence of outliers.

\paragraph{Variability.}
Variability captures the degree to which a set of data differs from the central tendency and is sometimes also referred to as the dispersion or spread of the data \citep{weisberg1992}. Common measurements of the variability are the variance or standard deviation as well as their robust counter parts. In time series analyses, variability is conceptually important because information about the distribution and diversity of the data has been found to be indicative of worse psychological states \citep{myin-germeys2018, helmich2021}. Person-level differences of ESM measurements have, for example, been associated with higher levels of psycho-pathological recurrences among depression patients \citep{timm2017}. As such, psychological researchers and practitioners are often empirically interested in between-person differences in variability. Researchers on polarization and radicalization might for example ask: "Are people settled in their attitudes towards migrants or do they vary across the measurement period?"

For our illustration data, we chose the \textit{Median Absolute Deviation} ($MAD$) to gauge the variability within our time series data. This choice is motivated by the robustness of MAD, particularly its resilience to the effects of non-normal distributions and outliers, which can significantly skew traditional variability measures like the standard deviation \citep{weisberg1992}. For a given variable \( j \) and participant \( i \), the MAD is calculated by first determining the median (\( M \)) of the ordered time series \( \mathbf{X}_{ij} \) as outlined in \equatref{eq:median}. We then compute the absolute deviations of each observation in the time series \( X_{ij} \) from this median value. Specifically, for each time point \( t \), we calculate the absolute difference between \( x_{ijt} \) and the series median \( M(\mathbf{X}_{ij}) \). The MAD is then the median of these absolute deviations:

```{=tex}
\begin{subequations}\label{eq:mad}
    \begin{align} 
      MAD(X_{ij}) &= M(\left| x_{ijt} - M(\mathbf{X}_{ij}) \right|) \label{eq:mad_general} \\
                  &= M(\{ |x_{ij1} - M(\mathbf{X}_{ij})|, |x_{ij2} - M(\mathbf{X}_{ij})|, \ldots, |x_{ijn} - M(\mathbf{X}_{ij})| \}) \label{eq:mad_detailed}
    \end{align}
\end{subequations}
```

The calculation of MAD focuses on the magnitudes of deviations, ensuring it provides a robust measure of dispersion that reflects the inherent variability in the time series data.

\paragraph{Instability.}
Instability captures the average change between two consecutive measurements \citep{ebner-priemer2009, jahng2008}. While instability is conceptually related to the variability feature, variability does not take into account temporal dependency, whereas instability looks at the 'jumpy-ness' of the data over time. In other words, variability reflects the range or diversity of values in the un-ordered time series data, while instability reflects the fluctuation or inconsistency in a time series data over time \citep{trull2008, houben2015, koval2013}. For example, if a person has rapid and extreme changes in mood their mood is highly unstable, while if a person's mood responses span a wide range over the entire study period, their mood is highly variable \citep[note that this does not need to be rapidly changing or instable, e.g., when there is linear increase over time; also see][]{jahng2008}. Within psychological time series, instability measurements have especially been important in the research of borderline personality disorder \citep{trull2008} and suicidality \citep{kivela2022}, but also in understanding early warning signals more generally \citep{wichers2019}. Conceptually, the instability feature, thus, relates to a broad range of research questions, including: "What is the nature of the identification changes in those who start working in a new country?" or "Do strong daily fluctuations in self-esteem reflect the process of identity formation in adolescents?"

For our data we chose the \textit{mean absolute change} \citep[$MAC$; e.g.,][]{ebner-priemer2009, barandas2020}, which looks at the average absolute difference of two consecutive measurements $x$ at time points $t$ and $t-1$, for each time series $X$ of participant $i$ and variable $j$.

```{=tex}
\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=2, \ldots, t}\left|x_{t}-x_{t-1}\right|
\end{equation}
```

Another common measurement of instability is the \textit{Mean of the Squared Successive Differences} (\(MSSD\)), which is often preferred where differences in magnitude are more important than the frequency of those changes, for example, when big shifts in time series are considered more impactful or when outliers are meaningful and need to be taken into account \citep{chatfield2003, bos2019}. For psychological ESM data, some research suggests that amplitude and frequency might predict different health outcomes and can be investigated jointly \citep{wang2012a, jahng2008}.

\paragraph{Temporal dependence.}
Temporal dependence in time series data refers to the degree to which a time series is influenced by its past values, exhibiting patterns of behavior that may be regular over different time scales \citep{dmello2021}. In the context of psychological time series, an important aspect of temporal dependence is \textit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}. If inertia is high a development tends to stay in a certain state. Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of under-reactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}. In a similar vein, high inertia in negative affect at baseline was predictive of the initial onset of depression \citep{kuppens2012}. Conceptually, inertia is more broadly connected to research questions such as: "Do patients stay in a negative mood for several measurements?" or "Do migrants stay with their language practice for several days at a time?"

For our illustration case, we chose the commonly used autocorrelation or autoregression with a lag-1 to capture the inertia. High autocorrelation values can indicate high levels of inertia, while low autocorrelation values may indicate a more unpredictable or volatile time series \citep{dejonckheere2019}. The lag--1 autocorrelation $r_{ij,1}$ looks at the average correlation between a measurement $x$ and the preceding measurement $x_{t-1}$ for the time series $X$ of participant $i$ and variable $j$ with $n$ measurements.

```{=tex}
\begin{equation} \label{eq:ar}
  r_{ij,1} = \frac{\sum_{t=2}^{n}(x_{ijt}-\overline{x}_{ij})(x_{ij,t-1}-\overline{x}_{ij})}{\sum_{t=1}^{n}(x_{ijt}-\overline{x}_{ij})^2}
\end{equation}
```

Where $\overline{x}_{ij}$ is the mean of the time series $x_{ij}$, calculated as:

```{=tex}
\begin{equation} \label{eq:mean_for_ar1}
  \overline{x}_{ij} = \frac{1}{n} \sum_{t=1}^{n} x_{ijt}
\end{equation}
```

While inertia captures the simplest case of temporal dynamics, lag-1, we acknowledge that temporal dependence in psychological time series may also exhibit more complex relationships, including higher lagged auto correlations or cyclical relationships (fourier coefficients, or continuous wavelet transforms are often used to capture such relationships).

\paragraph{Linear trend.}
In non-stationary time series, a linear trend can be observed when there is a consistent increase or decrease in the data over time \citep{nyblom1986}. For psychological time series, researchers have, for example, pointed out the importance of linear trends in interpersonal communications \citep{vasileiadou2014}, and emotion dynamics \citep{oravecz2016}. Theoretically, linear trends are often considered the simplest way of assessing whether a psychological theory of change is appropriate \citep{gottman1969}. In empirical practice, linear trends are, thus, commonly exemplified by research questions such as "Do patient symptoms improve consistently?" or "Does worker productivity decline continuously?"

For the variables in our illustration data set, we chose an overall linear regression slope to capture the linear trend. The regression slope $b_{ij}$ provides the average change from one time point $t$ to the next across all measurements $x$ of a time series $X$ of participant $i$ and variable $j$. The specific form of the OLS slope formula we provide below calculates $b_{ij}$ as the sum across all time points of the product of the deviation of time $t$ from its mean $\overline{t}$ and the deviation of $x_{ij}$ from its mean $\overline{x}_{ij}$ at each time point, divided by the sum across all time points of the square of the deviation of time from its mean ($\sum(t-\overline{t})^2$). Intuitively, the formula captures the rate of change of variable $x_{ij}$ with respect to time. This slope will indicate how the variable $x_{ij}$ changes over time, controlling for its mean value and the mean of time. If the slope is positive, $x_{ij}$ increases over time; if it's negative, $x_{ij}$ decreases over time.

```{=tex}
\begin{equation} \label{eq:lin}
  b_{ij} = \frac{\sum(t-\overline{t})(x_{ijt}-\overline{x}_{ij})}{\sum(t-\overline{t})^2}
\end{equation}
```

\paragraph{Nonlinearity.}
Changes in psychology are not always linear, instead, nonlinearity is a common feature of psychological time series \citep{hayes2007}. As an example, episodic disorders, such as depression, are often best described as non-linear systems \citep{hosenfeld2015}. Similarly, patients in recovery from depression showed sudden changes in the improvement of depression \citep{helmich2020a}. But also substance abuse \citep{boker1998} or attitude changes rarely develop linearly \citep{vandermaas2003}. Conceptually, researchers might have research questions about the type of the development: "Is the development of well-being a nonlinear process?" as well as the shape and structure of the development: "How many spikes in well-being did a migrant experience?"

We summarized the nonlinear trend with the \textit{estimated degrees of freedom} of an empty GAM spline model. The $edf$ summarizes the \textit{wiggliness} of a spline trend line \citep{wood2017, bringmann2017}. The degrees of freedom of a spline model are primarily determined by the number of knots and the order of the spline. For instance, a cubic spline with $k$ knots has $k$+3 degrees of freedom \citep{faraway2016, haslbeck2021a, castro-alvarez2024}. However, in a penalized spline framework, which is commonly used for GAMs, the effective degrees of freedom can be less than $k$+3. This is because the model employs a smoothing parameter to control the trade-off between the complexity (flexibility) of the model and its fit to the data, thereby penalizing overly complex models and potentially reducing the effective degrees of freedom \citep{marx1998}. Intuitively then an edf of 1 would be equivalent to a linear relationship (i.e., one linear slope parameter), whereas a higher edf (particularly an edf > 2) is indicative of a non-linear trend. The estimated degrees of freedom are commonly based on a concept called 'effective degrees of freedom' and can be represented as the trace $tr$ (i.e., the sum of the diagonal elements) of the smoother matrix $S$, a symmetric matrix that maps from the raw data to the smooth estimates \citep{wood2017}.

```{=tex}
\begin{equation} \label{eq:df}
  edf = tr(S)
\end{equation}
```

```{r}
#| include: False

# Effective degrees of freedom for individual smooth terms are obtained by summing the Fii values corresponding to the coefficients βi of the smooth term (e.g., if the coefficients of a particular smooth are β5, . . . , β11 then the smooth’s EDF is $\sum_{i=5}^{11} = F_{ii}).
# 
# An alternative approach to understanding the edf is to approximate the number of parametric and non-parametric (or smooth) terms $p$ in the penalized GAM. Both formulas are measuring the same concept (the model's complexity or flexibility) and both account for the penalization introduced to control overfitting in GAMs \citep{hastie1986}.
# 
# ```{=tex}
# \begin{equation} \label{eq:edf}
#   edf \approx \sum p
# \end{equation}
# ```
```

\paragraph{Additional considerations.}
Beyond our main features of interest, we also extracted the participant's number of completed ESM measurements to ensure that the clusters are comparable in that regard (i.e., to exclude spurious explanations for the cluster assignments). After the feature extraction, we found that about `r format(round(pMissFull$pmiss %>% filter(set == "All") %>% select(pMissTotal) %>% pull, 2), nsmall=2)`% of the extracted features are missing across the `r ncol(featFull$features %>% select(-ends_with(featureExclusion)))-1` features per participant. This might, for example, happen if participants do not have two subsequent measurements with outgroup interactions, so that an autocorrelation with lag-1 cannot be calculated for the contact-specific variables. The small number of missing values indicates that the feature-based approach indeed largely avoids the structural missingness issue. However, even the few missing values can be an issue for some feature reduction or feature clustering algorithms. We, thus, impute missing feature values via predictive mean matching (PMM) with the MICE package in R, employing a single imputation and specifying a maximum of 50 iterations and a fixed seed for convergence and reproducibility \citep[][]{buuren2011}. We chose PMM for its ability to preserve the original data distribution without assuming normality and robustly handling multiple data types \citep{vanbuuren2006}. Note again that with this procedure we only need to impute an extremely small number of missing values as most feature calculations can use the available data instead.